### 2024-11-19 MS&E 226_Project Part 1_Code.Rmd
Add files via upload

---
title: "MS&E 226: Project Part 1 Code"
author: "Antra Nakhasi & Sydney Loh"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, warning = FALSE)
knitr::opts_chunk$set(tidy = TRUE, tidy.opts = list(width.cutoff = 60))

options(width = 60) # Set width for knitting code chunks

rm(list = ls()) # Clear environment

# Load necessary libraries
library(tidyverse)
library(dplyr)
library(caret)
library(kknn)
library(reshape2)
library(class)
library(glmnet)
library(e1071)
library(rpart)
library(tibble)
library(randomForest)   
library(xgboost)  
library(quantreg)
library(kernlab)

set.seed(123) # Setting seed for reproducibility
```

```{r Pre Process, results='hide'}

# load data
df <- read.csv("/Users/sydneyloh/OneDrive/Documents/Stanford/MS&E 226/Term Project/electricity.csv")

## pre process data

# Convert Revenues.Retail to Revenue.Retail to match singular use of "Revenue" in all other column names
colnames(df)[colnames(df) == "Revenues.Retail"] <- "Revenue.Retail"

# Create binary outcome variable from Uses.Consumed (0 if = 0, 1 otherwise)
df$Uses.Consumed.Binary <- ifelse(df$Uses.Consumed == 0, 0, 1)

# Remove Retail.Total.Sales and Retail.Total.Customers columns (b/c not linearly independent, keep more granular columns)
df <- df[, !names(df) %in% c("Retail.Total.Customers", "Retail.Total.Sales", "Sources.Total", "Uses.Total", 
                             "Retail.Total.Revenue", "Retail.Total.Sales", "Retail.Total.Customers")]

# Convert revenue breakdown columns, Revenue.______, into a categorical variable listing which stream had the greatest share of the total
revenue_cols <- c("Retail.Residential.Revenue", "Retail.Commercial.Revenue", 
                  "Retail.Industrial.Revenue", "Retail.Transportation.Revenue")

# Find the sector with the maximum revenue and remove the revenue columns
df <- df %>%
  rowwise() %>%
  mutate(
    Max.Revenue.Sector = c("Residential", "Commercial", "Industrial", "Transportation")[which.max(c_across(all_of(revenue_cols)))]
  ) %>%
  ungroup() %>%
  select(-all_of(revenue_cols))

northeast_states <- c("CT", "ME", "MA", "NH", "RI", "VT", "NJ", "NY", "PA", "DC")
midwest_states <- c("IL", "IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD")
south_states <- c("DE", "FL", "GA", "MD", "NC", "SC", "VA", "WV", "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX")
west_states <- c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AK", "CA", "HI", "OR", "WA")

# Assign each state to its corresponding region
df$Utility.Region <- case_when(
  df$Utility.State %in% northeast_states ~ "Northeast",
  df$Utility.State %in% midwest_states ~ "Midwest",
  df$Utility.State %in% south_states ~ "South",
  df$Utility.State %in% west_states ~ "West",
  df$Utility.State == "CN" ~ "Canada",
  TRUE ~ NA_character_ 
)

# Found one NA value where Utility.State was a blank string "" 
# Omit since it was one case at random 
df <- df[-which(is.na(df$Utility.Region)), ]

```

```{r Create Holdout Set, results='hide'}
# Total number of rows in the dataset
n <- nrow(df)

# Sample 20% of the data for the holdout set
holdout_indices <- sample(1:n, size = floor(0.2 * n))

# Create holdout and training sets
holdout <- df[holdout_indices, ]
df <- df[-holdout_indices, ]

# Display the sizes of the datasets
cat("Holdout Set Size:", nrow(holdout), "\n")
cat("Training Set Size:", nrow(df), "\n")
```

```{r Correlations, fig.show='hide', results='hide'}

# Correlation the continuous outcome: Hours Worked Per Week

# Select relevant columns for analysis
columns_of_interest <- names(df)[!names(df) %in% c("Utility.Number", "Utility.Name", "Utility.State")] # use a different visualization since 1-hot encoding would lead to 50 different states

# Filter data to include only columns of interest
data_subset <- df %>% select(all_of(columns_of_interest))

# Separate numeric and categorical columns
numeric_columns <- names(data_subset)[sapply(data_subset, is.numeric)]
categorical_columns <- setdiff(columns_of_interest, numeric_columns)

# One-hot encode categorical variables
dummies <- dummyVars(" ~ .", data = data_subset[categorical_columns])
encoded_categorical_data <- predict(dummies, newdata = data_subset) %>% as.data.frame()

# Combine numeric data with one-hot encoded categorical data
encoded_data <- cbind(data_subset[numeric_columns], encoded_categorical_data)

# Calculate all correlations
correlations <- cor(encoded_data, use = "complete.obs")

# Calculate correlations with "Hours_Worked_Per_Week"
correlations_with_revenue <- correlations["Revenue.Total",]

# Sort and display correlations
sorted_correlations_revenue <- sort(correlations_with_revenue, decreasing = TRUE)
# Create a data frame of sorted correlations for better readability
sorted_correlations_revenue <- data.frame(
  Variable = names(sorted_correlations_revenue),
  Correlation_with_revenue = sorted_correlations_revenue
)

# Display the data frame in a readable format
print(sorted_correlations_revenue, row.names = FALSE)

top_correlated_vars <- head(sorted_correlations_revenue$Variable[-1], 5)
# Generate scatterplots for each of the top correlated variables with Revenue.Total
for (var in top_correlated_vars) {
  p <- ggplot(df, aes_string(x = var, y = "Revenue.Total")) +
    geom_point(alpha = 0.6, color = "blue") +
    labs(title = paste("Scatterplot of", var, "vs Revenue.Total"),
         x = var, y = "Revenue.Total") +
    theme_minimal() +
    geom_smooth(method = "lm", color = "red", se = FALSE)

  print(p)  # Explicitly print each plot
}

ggplot(sorted_correlations_revenue, aes(x = reorder(Variable, Correlation_with_revenue), 
                                             y = Correlation_with_revenue)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for better readability
  labs(title = "Correlation with Revenue.Total", 
       x = "Variables", 
       y = "Correlation Coefficient") +
  theme_minimal()

# Calculate correlations with binary outcome variable
correlations_with_uses_consumed <- correlations["Uses.Consumed.Binary",]

# Sort and display correlations
sorted_correlations_uses_consumed <- sort(correlations_with_uses_consumed, decreasing = TRUE)

# Create a data frame of sorted correlations for better readability
sorted_correlations_uses_consumed <- data.frame(
  Variable = names(sorted_correlations_uses_consumed),
  Correlation_with_Uses_Consumed = sorted_correlations_uses_consumed
)

# Display the data frame in a readable format
print(sorted_correlations_uses_consumed, row.names = FALSE)

ggplot(sorted_correlations_uses_consumed, aes(x = reorder(Variable, Correlation_with_Uses_Consumed), 
                                             y = Correlation_with_Uses_Consumed)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for better readability
  labs(title = "Correlation with Revenue.Total", 
       x = "Variables", 
       y = "Correlation Coefficient") +
  theme_minimal()

```

```{r correlation of covariates with each other, fig.show='hide', results='hide'}

# Create a list of numeric columns, not including the binary outcome variable 
cont_cov_columns <- numeric_columns[numeric_columns != "Uses.Consumed.Binary"]

# Calculate correlations for continuous variables
correlations_cont <- cor(data_subset[cont_cov_columns], use = "complete.obs")

# Melt the correlation matrix for ggplot2
cor<- melt(correlations_cont)

# Create the heatmap
ggplot(data = cor, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 10, hjust = 1)) +
  coord_fixed()

```

```{r 5f subgroups in data, fig.show='hide', results='hide'}

# Calculate means for utility.state
state_means <- df %>%
  summarise(across(-c(Utility.Number, Utility.Name, Utility.Type, Max.Revenue.Sector, Uses.Consumed.Binary, Utility.Region),
                   ~ round(mean(.x, na.rm = TRUE), 5)),
            .by = Utility.State)

# Calculate means for utility.type
type_means <- df %>%
  summarise(across(-c(Utility.Number, Utility.Name, Utility.State, Max.Revenue.Sector, Uses.Consumed.Binary, Utility.Region),
                   ~ round(mean(.x, na.rm = TRUE), 5)),
            .by = Utility.Type)

region_means <- df %>%
  summarise(across(-c(Utility.Number, Utility.Name, Utility.State, Max.Revenue.Sector, Uses.Consumed.Binary, Utility.Type),
                   ~ round(mean(.x, na.rm = TRUE), 5)),
            .by = Utility.Region)

View(state_means)
View(type_means)
View(region_means)
variables_to_plot <- names(state_means)[-1]  # Exclude Utility.State
plot_list <- list()

# Loop through each variable and create a plot
for (var in variables_to_plot) {
  p <- ggplot(state_means, aes_string(x = "Utility.State", y = var)) +
    geom_bar(stat = "identity", fill = "green") +
    coord_flip() +
    labs(title = paste("Mean", var, "by Utility State"),
         x = "Utility State",
         y = paste("Mean", var)) +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 6, angle = 0),
          axis.title.y = element_text(size = 10),
          axis.title.x = element_text(size = 10),
          plot.title = element_text(size = 12, face = "bold"))

  plot_list[[var]] <- p
}

for (plot in plot_list) {
  print(plot)
}
# Create plots for utility type means
variables_to_plot_type <- names(type_means)[-1]  # Exclude Utility.Type
plot_list_type <- list()

# Loop through each variable and create a plot for utility type
for (var in variables_to_plot_type) {
  p <- ggplot(type_means, aes_string(x = "Utility.Type", y = var)) +
    geom_bar(stat = "identity", fill = "coral") +
    coord_flip() +
    labs(title = paste("Mean", var, "by Utility Type"),
         x = "Utility Type",
         y = paste("Mean", var)) +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 6, angle = 0),
          axis.title.y = element_text(size = 10),
          axis.title.x = element_text(size = 10),
          plot.title = element_text(size = 12, face = "bold"))

  plot_list_type[[var]] <- p
}

for (plot in plot_list_type) {
  print(plot)
}
# Define the list of variables to plot, excluding the grouping column (Utility.Region)
variables_to_plot <- names(region_means)[-1]

# Initialize an empty list to store the plots
plot_list <- list()

# Loop through each variable in region_means and create a plot
for (var in variables_to_plot) {
  p <- ggplot(region_means, aes_string(x = "Utility.Region", y = var)) +
    geom_bar(stat = "identity", fill = "green") +
    coord_flip() +
    labs(title = paste("Mean", var, "by Utility Region"),
         x = "Utility Region",
         y = paste("Mean", var)) +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 6, angle = 0),
          axis.title.y = element_text(size = 10),
          axis.title.x = element_text(size = 10),
          plot.title = element_text(size = 12, face = "bold"))

  # Add each plot to the list
  plot_list[[var]] <- p
}

# Print each plot for region_means
for (plot in plot_list) {
  print(plot)
}

```

# 6: Regression

```{r regression}
# set seed
set.seed(123)

### Choosing prediction error metric
# Calculate summary statistics for revenue data
mean_revenue <- mean(df$Revenue.Total)
median_revenue <- median(df$Revenue.Total)
std_dev_revenue <- sd(df$Revenue.Total)
iqr_revenue <- IQR(df$Revenue.Total)

# Print summary statistics
print(paste("Mean Revenue:", mean_revenue))
print(paste("Median Revenue:", median_revenue))
print(paste("Standard Deviation:", std_dev_revenue))
print(paste("IQR:", iqr_revenue))

columns_to_exclude <- c("Utility.Name", "Utility.Number", "Utility.State", 
                         "Revenue.Retail", "Revenue.Delivery", 
                         "Revenue.Resale", "Revenue.Adjustments", 
                         "Revenue.Transmission", "Revenue.Other", "Uses.Consumed") #Revenue subparts are linearly dependent on total revenue

# Create a new data frame without the excluded columns
# df_pred <- df[, !names(df) %in% columns_to_exclude]

# Define the columns to include in df_pred
column_names <- c(
  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Sources.Purchased",
  "Sources.Other", "Uses.Retail", "Uses.Resale", "Uses.No.Charge", "Uses.Consumed", "Uses.Losses",
  "Retail.Residential.Customers", "Retail.Commercial.Customers", "Retail.Industrial.Customers",
  "Retail.Transportation.Customers", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type"
)

# Create df_pred with only the specified columns from df
df_pred <- df[, column_names]

# # Display df_pred
# df_pred

### Applying data transformations
# Applying log transformation to the Revenue.Total column
df_pred$Log_Revenue_Total <- log1p(df_pred$Revenue.Total)

# Histogram
ggplot(df_pred, aes(x = Revenue.Total)) + 
  geom_histogram(binwidth = 5000, fill = "skyblue", color = "black", alpha = 0.7) + 
  labs(title = "Histogram of Revenue Total", x = "Revenue Total", y = "Frequency") +
  theme_minimal()

# Histogram of log-transformed Revenue.Total (after scaling)
ggplot(df_pred, aes(x = Log_Revenue_Total)) + 
  geom_histogram(binwidth = 0.1, fill = "lightgreen", color = "black", alpha = 0.7) + 
  labs(title = "Histogram of Log-transformed Revenue Total", 
       x = "Log(Revenue Total)", 
       y = "Frequency") +
  theme_minimal()
### Baseline Model
# Fit OLS Model (Linear Regression) without cross-validation
ols_model <- lm(Log_Revenue_Total ~ ., 
                data = df_pred)  # Fit the model on the entire dataset

# Print the summary of the OLS model
summary(ols_model)

# Get the residuals to calculate RMSE
predictions <- predict(ols_model, newdata = df_pred)
residuals <- df_pred$Log_Revenue_Total - predictions

# Calculate RMSE (Root Mean Squared Error)
rmse <- sqrt(mean(residuals^2))
cat(paste("OLS Model RMSE:", rmse, "\n"))

### Cross Validation
# Set up cross-validation control with 10 folds
train_control <- trainControl(method = "cv", 
                              number = 10, 
                              summaryFunction = defaultSummary,
                              search = "grid",
                              savePredictions = "all")
### Ridge Regression
ridge_model <- train(Log_Revenue_Total ~ ., 
                     data = df_pred, 
                     method = "glmnet",  # glmnet is used for regularized regression
                     trControl = train_control,
                     metric = "RMSE", 
                     preProcess = "scale",  # Standardize the data
                     tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 1, length = 100))) # alpha=0 for Ridge

# Print CV error 
cat("Ridge Regression CV RMSE for each fold:\n")
print(ridge_model$resample$RMSE)

ridge_model_rmse <- min(ridge_model$results$RMSE)
cat(paste("Ridge Regression RMSE:", ridge_model_rmse, "\n"))

final_lambda_ridge <- ridge_model$bestTune$lambda
cat(paste("Final lambda value for Ridge regression:", final_lambda_ridge, "\n"))
### Lasso Regression
# Fit Lasso regression using glmnet with tuning of lambda and data standardization
lasso_model <- train(Log_Revenue_Total ~ ., 
                     data = df_pred, 
                     method = "glmnet",  # glmnet for regularized regression
                     trControl = train_control,
                     metric = "RMSE",  # RMSE for model evaluation
                     preProcess = "scale",  # Standardize the data
                     tuneGrid = expand.grid(alpha = 1, lambda = seq(0, 1, length = 100)))  # Lasso: alpha = 1

# Print CV error (RMSE for each fold)
cat("Lasso Regression CV RMSE for each fold:\n")
print(lasso_model$resample$RMSE)

# Extract and print the best RMSE for Lasso Regression
lasso_model_rmse <- min(lasso_model$results$RMSE)
cat(paste("Lasso Regression Best RMSE:", lasso_model_rmse, "\n"))

# Extract and print the best lambda value for Lasso regression
final_lambda_lasso <- lasso_model$bestTune$lambda
cat(paste("Final lambda value for Lasso regression:", final_lambda_lasso, "\n"))
### Random Forest
# Fit a Random Forest model
rf_model <- train(Log_Revenue_Total ~ ., 
                  data = df_pred, 
                  method = "rf",  # Random Forest regression
                  trControl = train_control,
                  metric = "RMSE",
                  tuneGrid = expand.grid(mtry = sqrt(ncol(df_pred)))) # Hyperparameter tuning for mtry

# Print CV error
cat("Random Forest CV RMSE for each fold:\n")
print(rf_model$resample$RMSE)

rf_model_rmse <- min(rf_model$results$RMSE)
cat(paste("Random Forest RMSE:", rf_model_rmse, "\n"))

```

# 7: Classification

```{r clean data for building classifiers}

# Select specific columns from the dataframe based on positive correlation > 0.1 or negative correlation <-0.1
df_clean <- df[, c("Uses.Consumed.Binary", # Binary outcome variable
                   "Utility.Type", # Positively or negatively correlated with outcome, depending on factor level 
                   "Demand.Winter.Peak", # Positively correlated with outcome
                   "Demand.Summer.Peak", # Positively correlated with outcome
                   "Max.Revenue.Sector", # Negatively correlated with outcome
                   "Revenue.Retail", # Positively correlated with outcome
                   grep("^Retail\\.", names(df), value = TRUE))] # Positively correlated with outcome

# Loop through each column in the dataset and apply the factor transformation to each non-numeric (categorical) variable
df_clean[] <- lapply(df_clean, function(x) {
  if (!is.numeric(x)) {
    factor(x, levels = unique(x))
  } else {
    x
  }
})

df_clean$Uses.Consumed.Binary <- factor(df_clean$Uses.Consumed.Binary, levels = c("0", "1"))
levels(df_clean$Uses.Consumed.Binary) <- make.names(levels(df_clean$Uses.Consumed.Binary))

# Create a copy of the original dataframe to store the scaled data
df_clean_scaled <- df_clean

# Identify numeric columns, excluding 'Uses.Consumed.Binary'
numeric_cols <- sapply(df_clean, is.numeric) & names(df_clean) != "Uses.Consumed.Binary"

# Apply scale() to the selected numeric columns
df_clean_scaled[numeric_cols] <- scale(df_clean[numeric_cols])

```

```{r Classifier 1: K-NN}

# Split the data into features and target variable
X <- df_clean_scaled[, !names(df_clean) %in% "Uses.Consumed.Binary"] # continuous outcome variable and linear combination of other columns
# Step 1: Prepare the data
# Identify numeric and categorical columns
numeric_cols <- sapply(X, is.numeric)
categorical_cols <- sapply(X, function(x) is.factor(x) | is.character(x))

# Create dummy variables for categorical features
dummy <- dummyVars(" ~ .", data = X[, categorical_cols])
categorical_dummy <- predict(dummy, newdata = X[, categorical_cols])

# Combine numeric and dummy-coded categorical variables
X <- cbind(X[, numeric_cols], categorical_dummy)

# Ensure the target variable is a factor
y <- df_clean_scaled$Uses.Consumed.Binary

# Step 2: Define a function to find the best k using cross-validation

# Define a custom summary function for 0-1 loss
zero_one_loss <- function(data, lev = NULL, model = NULL) {
  loss <- mean(data$obs != data$pred)
  c(loss = loss)
}

# Find_best_k function
find_best_k <- function(X, y, k_range, distance_metric) {
  set.seed(123)

  # Define custom summary function for 0-1 loss
  zero_one_loss <- function(data, lev = NULL, model = NULL) {
    loss <- mean(data$obs != data$pred)
    c(loss = loss)
  }

  control <- trainControl(method = "cv",
                          number = 10,
                          summaryFunction = zero_one_loss)

  grid <- expand.grid(k = k_range)

  knn_model <- train(x = X, y = y,
                     method = "knn",
                     tuneGrid = grid,
                     trControl = control,
                     metric = "loss",
                     maximize = FALSE)  # We want to minimize the loss

  best_k <- knn_model$bestTune$k
  best_loss <- min(knn_model$results$loss)

  return(list(best_k = best_k, mean_0_1_loss = best_loss))
}

# Step 3: Find the best k
k_range <- seq(1, 20, by = 2)
best_k <- find_best_k(X, y, k_range)

# Step 4: Train the final k-NN model with the best k
final_model <- knn(train = X, test = X, cl = y, k = best_k[[1]])

# Save mean 0-1 loss for best k
Err_CV_knn <- best_k[[2]]

# Print results
print(paste("Best k:", best_k[[1]]))
print(paste("0-1 Loss:", round(Err_CV_knn, 4)))

```

```{r Classfier 2: Logistic Regression with Feature Selection}

# Create model formula
formula <- as.formula("Uses.Consumed.Binary ~ Utility.Type + Max.Revenue.Sector + Demand.Winter.Peak + Demand.Summer.Peak + Revenue.Retail + Revenue.Retail:Max.Revenue.Sector")

# Set up cross-validation for classification
ctrl <- trainControl(
  method = "cv",
  number = 10,
  classProbs = TRUE,
  summaryFunction = defaultSummary
)

# Define grid of lambda values to tune for regularization
tune_grid <- expand.grid(
  alpha = 0,       # Set alpha = 0 for ridge regression (L2 regularization)
  lambda = seq(0.001, 0.1, by = 0.01) # Define a sequence of lambda values
)

# Train the logistic regression model with cross-validation and lambda tuning
model_cv <- train(
  formula,
  data = df_clean_scaled,
  method = "glmnet",
  family = "binomial",
  trControl = ctrl,
  metric = "Accuracy",
  tuneGrid = tune_grid
)

# Calculate 0-1 loss for each fold
zero_one_loss <- 1 - model_cv$resample$Accuracy

# Calculate mean 0-1 loss across folds
Err_CV_logr <- mean(zero_one_loss)

# Train final model on all data with the best lambda
final_model <- glmnet(
  as.matrix(df_clean_scaled[, all.vars(formula)[-1]]), 
  df_clean_scaled$Uses.Consumed.Binary, 
  alpha = 0,
  lambda = model_cv$bestTune$lambda,
  family = "binomial"
)

# Print results
print(paste("0-1 Loss:", round(Err_CV_logr, 4)))
print(paste("Best lambda:", model_cv$bestTune$lambda))

```

```{r Classifier 3: Decision Tree}

set.seed(123)  # For reproducibility

# Create model formula
formula <- as.formula("Uses.Consumed.Binary ~ Utility.Type + Max.Revenue.Sector + Demand.Winter.Peak + Demand.Summer.Peak + Revenue.Retail + Revenue.Retail:Max.Revenue.Sector")

# Define the hyperparameter grid
hyper_grid <- expand.grid(cp = seq(0.01, 0.1, by = 0.01))

# Set up 10-fold cross-validation with Accuracy as the selection metric
ctrl <- trainControl(
  method = "cv",
  number = 10
)

# Perform grid search with cross-validation
model_grid <- train(
  formula,
  data = df_clean_scaled,
  method = "rpart",
  trControl = ctrl,
  tuneGrid = hyper_grid,
  metric = "Accuracy"
)

# Get the best model
best_model <- model_grid$finalModel

# Calculate the average cross-validation error (1 - Accuracy)
Err_CV_dt <- 1 - max(model_grid$results$Accuracy)

# Display the average cross-validation error
print(paste("0-1 Loss:", round(Err_CV_dt, 4)))
# Display the best model parameters
print(paste("cp of chosen model:", model_grid$bestTune))

```

```{r compare classifiers, include=TRUE, echo=FALSE, results='asis'}

Err_baseline <- round(sum(df$Uses.Consumed.Binary)/nrow(df), 4)

Errs <- c(Baseline = Err_baseline, Knn = round(Err_CV_knn, 4), Logr = round(Err_CV_logr, 4), DT = round(Err_CV_dt, 4))

# Convert to a tibble
Errs_table <- tibble(Model = c("Baseline", "K-Nearest Neighbors", "Logistic Regression (Ridge)", "Decision Tree"), Error = sprintf("%.4f", Errs))

# Print the table
knitr::kable(Errs_table, col.names = c("Model", "Prediction Error (Mean 0-1 Loss)"))

print(Errs_table)
```

### 2024-11-19 MS&E 226_Project Part 1_Code.Rmd
comments added

@@
 
+# 2: Data Pre-processing
+
 ```{r Pre Process, results='hide'}
@@
 # load data
-df <- read.csv("/Users/sydneyloh/OneDrive/Documents/Stanford/MS&E 226/Term Project/electricity.csv")
+df <- read.csv("electricity.csv")
 
@@
 
+# 3: Hold Out Set
+
 ```{r Create Holdout Set, results='hide'}
@@
 
-```{r Correlations, fig.show='hide', results='hide'}
+# 5: Data Exploration
+
+## Correlations
+
+```{r Correlations with Revenue.Total, fig.show='hide', results='hide'}
 
@@
   theme_minimal()
+```
 
+```{r Correlations with Uses.Consumed.Binary, fig.show='hide', results='hide'}
 # Calculate correlations with binary outcome variable
@@
 
-```{r 5f subgroups in data, fig.show='hide', results='hide'}
+## Subgroups in data
+
+```{r means across categories, fig.show='hide', results='hide'}
 
@@
 plot_list <- list()
+```
 
+```{r Utility.State plots, fig.show='hide', results='hide'}
 # Loop through each variable and create a plot
@@
 }
+
+```
+
+```{r Utility.Type plots, fig.show='hide', results='hide'}
 # Create plots for utility type means
@@
 }
+
+```
+
+```{r Utility.Region Plots, fig.show='hide', results='hide'}
 # Define the list of variables to plot, excluding the grouping column (Utility.Region)
@@
 }
-
 ```
@@
 
-```{r regression}
+## Data Transformations
+
+```{r data transformations}
 # set seed
@@
 print(paste("IQR:", iqr_revenue))
-
-columns_to_exclude <- c("Utility.Name", "Utility.Number", "Utility.State", 
-                         "Revenue.Retail", "Revenue.Delivery", 
-                         "Revenue.Resale", "Revenue.Adjustments", 
-                         "Revenue.Transmission", "Revenue.Other", "Uses.Consumed") #Revenue subparts are linearly dependent on total revenue
-
-# Create a new data frame without the excluded columns
-# df_pred <- df[, !names(df) %in% columns_to_exclude]
 
@@
   theme_minimal()
+
+```
+
+## Baseline Model
+
+```{r Baseline Model}
 ### Baseline Model
@@
 cat(paste("OLS Model RMSE:", rmse, "\n"))
+```
 
+## Cross Validation setup
+
+```{r cross validation}
 ### Cross Validation
@@
                               savePredictions = "all")
+```
+
+## Model 1: Ridge Regression
+
+```{r Model 1: Ridge Regression}
 ### Ridge Regression
@@
 cat(paste("Final lambda value for Ridge regression:", final_lambda_ridge, "\n"))
+
+```
+
+## Model 2: Lasso Regression
+
+```{r Model 2: Lasso Regression}
 ### Lasso Regression
@@
 cat(paste("Final lambda value for Lasso regression:", final_lambda_lasso, "\n"))
+
+```
+
+## Model 3: Random Forest
+
+```{r Model 3: Random Forest}
 ### Random Forest
@@
 cat(paste("Random Forest RMSE:", rf_model_rmse, "\n"))
-
 ```
@@
 # 7: Classification
+
+## Data Cleaning
 
@@
 ```
+
+## Classifier 1: K-NN
 
@@
 
+## Classifier 2: Logistic Regression with Feature Selection
+
 ```{r Classfier 2: Logistic Regression with Feature Selection}
@@
 
+## Classifier 3: Decision Tree
+
 ```{r Classifier 3: Decision Tree}
@@
 
+## Comparing Classifiers
+
 ```{r compare classifiers, include=TRUE, echo=FALSE, results='asis'}

### 2024-11-29 MS&E 226_Project Part 2_Code.Rmd
started part 2rmd file

--- 2024-11-19 MS&E 226_Project Part 1_Code.Rmd
@@
 # Identify numeric and categorical columns
-numeric_cols <- sapply(X, is.numeric)
-categorical_cols <- sapply(X, function(x) is.factor(x) | is.character(x))
+knn_numeric_cols <- sapply(X, is.numeric)
+knn_categorical_cols <- sapply(X, function(x) is.factor(x) | is.character(x))
 
 # Create dummy variables for categorical features
-dummy <- dummyVars(" ~ .", data = X[, categorical_cols])
-categorical_dummy <- predict(dummy, newdata = X[, categorical_cols])
+dummy <- dummyVars(" ~ .", data = X[, knn_categorical_cols])
+categorical_dummy <- predict(dummy, newdata = X[, knn_categorical_cols])
 
 # Combine numeric and dummy-coded categorical variables
-X <- cbind(X[, numeric_cols], categorical_dummy)
+X <- cbind(X[, knn_numeric_cols], categorical_dummy)
 
@@
 # Step 4: Train the final k-NN model with the best k
-final_model <- knn(train = X, test = X, cl = y, k = best_k[[1]])
+knn_class_final_model <- knn(train = X, test = X, cl = y, k = best_k[[1]])
 
@@
 # Create model formula
-formula <- as.formula("Uses.Consumed.Binary ~ Utility.Type + Max.Revenue.Sector + Demand.Winter.Peak + Demand.Summer.Peak + Revenue.Retail + Revenue.Retail:Max.Revenue.Sector")
+lr_formula <- as.formula("Uses.Consumed.Binary ~ Utility.Type + Max.Revenue.Sector + Demand.Winter.Peak + Demand.Summer.Peak + Revenue.Retail + Revenue.Retail:Max.Revenue.Sector")
 
@@
 model_cv <- train(
-  formula,
+  lr_formula,
   data = df_clean_scaled,
@@
 # Train final model on all data with the best lambda
-final_model <- glmnet(
-  as.matrix(df_clean_scaled[, all.vars(formula)[-1]]), 
+lr_class_final_model <- glmnet(
+  as.matrix(df_clean_scaled[, all.vars(lr_formula)[-1]]), 
   df_clean_scaled$Uses.Consumed.Binary, 
@@
 # Get the best model
-best_model <- model_grid$finalModel
+dt_class_final_model <- model_grid$finalModel
 
@@
 ```
+
+# Project Part 2
+
+## 1: Prediction on the holdout set
+
+### Classification
+
+#### a)
+
+In part 1, the logistic regression classifier had the best performance. This is the classifier that will be used for prediction on the holdout set.
+
+```{r Classification prediction with winning modeling strategy}
+# Select specific columns from the dataframe based on positive correlation > 0.1 or negative correlation <-0.1
+holdout_clean <- holdout[, c("Uses.Consumed.Binary", # Binary outcome variable
+                   "Utility.Type", # Positively or negatively correlated with outcome, depending on factor level 
+                   "Demand.Winter.Peak", # Positively correlated with outcome
+                   "Demand.Summer.Peak", # Positively correlated with outcome
+                   "Max.Revenue.Sector", # Negatively correlated with outcome
+                   "Revenue.Retail", # Positively correlated with outcome
+                   grep("^Retail\\.", names(df), value = TRUE))] # Positively correlated with outcome
+
+# Loop through each column in the dataset and apply the factor transformation to each non-numeric (categorical) variable
+holdout_clean[] <- lapply(holdout_clean, function(x) {
+  if (!is.numeric(x)) {
+    factor(x, levels = unique(x))
+  } else {
+    x
+  }
+})
+
+holdout_clean$Uses.Consumed.Binary <- factor(holdout_clean$Uses.Consumed.Binary, levels = c("0", "1"))
+levels(holdout_clean$Uses.Consumed.Binary) <- make.names(levels(holdout_clean$Uses.Consumed.Binary))
+
+# Scale numeric columns (use same scaling parameters as training data)
+holdout_clean_scaled <- holdout_clean
+numeric_cols_holdout <- sapply(holdout_clean, is.numeric)  # Identify numeric columns
+holdout_clean_scaled[numeric_cols_holdout] <- scale(
+  holdout_clean[numeric_cols_holdout], 
+  center = attr(scale(df_clean[numeric_cols]), "scaled:center"), 
+  scale = attr(scale(df_clean[numeric_cols]), "scaled:scale")
+)
+
+# Subset the holdout dataset to include only variables in the formula
+required_columns <- all.vars(formula)[-1]  # Exclude the response variable
+holdout_clean <- holdout[, required_columns, drop = FALSE]  # Keep only required columns
+
+# subset only covariates in the formula 
+formula_vars <- all.vars(lr_formula)
+
+# Subset the dataframe to include only the columns in the formula
+holdout_clean_scaled <- holdout_clean_scaled[, formula_vars, drop = FALSE]
+
+# Separate outcome and covarites
+
+Y_holdout <- holdout_clean_scaled$Uses.Consumed.Binary
+
+X_holdout <- holdout_clean_scaled[, colnames(holdout_clean_scaled) != "Uses.Consumed.Binary"]
+
+# Convert to matrix for glmnet predictions
+holdout_class_matrix <- as.matrix(X_holdout)
+
+# Make predictions (probabilities)
+predictions_prob <- predict(
+  lr_class_final_model, 
+  newx = holdout_class_matrix, 
+  s = model_cv$bestTune$lambda, 
+  type = "response"
+)
+
+# Convert probabilities to binary predictions
+predictions_binary <- ifelse(predictions_prob >= 0.5, "1", "0")
+
+# Calculate 0-1 loss
+true_labels <- holdout$Uses.Consumed.Binary  # True labels from holdout
+misclassified <- sum(predictions_binary != true_labels)
+zero_one_loss_holdout <- misclassified / length(true_labels)
+
+# Print results
+print(paste("0-1 Loss on holdout data:", round(zero_one_loss_holdout, 4)))
+
+```
+
+## 2: Inference
+
+#### (c)
+
+```{r fit model on the holdout set}
+
+# Define the columns to keep 
+column_names <- c(
+  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Sources.Purchased",
+  "Sources.Other", "Uses.Retail", "Uses.Resale", "Uses.No.Charge", "Uses.Consumed", "Uses.Losses",
+  "Retail.Residential.Customers", "Retail.Commercial.Customers", "Retail.Industrial.Customers",
+  "Retail.Transportation.Customers", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type"
+)
+
+# Create holdout_pred with only the specified columns from holdout
+holdout_pred <- holdout[, column_names]
+
+### Applying data transformations
+# Applying log transformation to the Revenue.Total column
+holdout_pred$Log_Revenue_Total <- log1p(holdout_pred$Revenue.Total)
+
+# Prepare the holdout dataset
+holdout_pred <- holdout[, column_names]
+holdout_pred$Log_Revenue_Total <- log(holdout_pred$Revenue.Total)
+
+# Prepare the data
+X <- model.matrix(Log_Revenue_Total ~ ., data = holdout_pred)[, -1]  # Design matrix (exclude intercept)
+y <- holdout_pred$Log_Revenue_Total  # Response variable
+
+# Prepare the input matrices for glmnet
+x <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]  # Remove intercept column
+y <- df_pred$Log_Revenue_Total
+
+# Fit the Ridge regression model with predefined lambda
+ridge_model <- glmnet(x, y, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
+
+# View model coefficients
+print(coef(ridge_model))
+
+```

### 2024-12-01 MS&E 226_Project Part 2_Code.Rmd
2c and 2d

@@
 library(kernlab)
+library(boot)
+library(ggplot2)
 
@@
 
-```{r fit model on the holdout set}
+```{r 2c Fit to test data Ridge}
 
-# Define the columns to keep 
+# manually refit final ridge model
+
+# Refit the model manually using glmnet
+X_train <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]
+y_train <- df_pred$Log_Revenue_Total
+
+final_ridge_model <- glmnet(X_train, y_train, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
+
+# Define the columns to include in df_pred
 column_names <- c(
@@
 
-# Create holdout_pred with only the specified columns from holdout
-holdout_pred <- holdout[, column_names]
+# save a copy of the holdout dataset to be transformed/cleaned 
+# Create df_test with only the specified columns from holdout
+df_test <- holdout[, column_names]
 
-### Applying data transformations
-# Applying log transformation to the Revenue.Total column
-holdout_pred$Log_Revenue_Total <- log1p(holdout_pred$Revenue.Total)
+# Log-transform the Revenue.Total column in the test set to match training
+df_test$Log_Revenue_Total <- log1p(df_test$Revenue.Total)
 
-# Prepare the holdout dataset
-holdout_pred <- holdout[, column_names]
-holdout_pred$Log_Revenue_Total <- log(holdout_pred$Revenue.Total)
+# Extract predictors (X) and response variable (y) for the test set
+X_test <- model.matrix(Log_Revenue_Total ~ ., data = df_test)[, -1]
+y_test <- df_test$Log_Revenue_Total
 
-# Prepare the data
-X <- model.matrix(Log_Revenue_Total ~ ., data = holdout_pred)[, -1]  # Design matrix (exclude intercept)
-y <- holdout_pred$Log_Revenue_Total  # Response variable
+# Refit the ridge model on the test set using the final lambda from training
+ridge_test <- glmnet(X_test, y_test, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
 
-# Prepare the input matrices for glmnet
-x <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]  # Remove intercept column
-y <- df_pred$Log_Revenue_Total
+# Extract coefficients from the test set
+ridge_test_coefs <- coef(ridge_test)
 
-# Fit the Ridge regression model with predefined lambda
-ridge_model <- glmnet(x, y, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
+# Extract coefficients from training and holdout models
+training_coefs <- as.matrix(coef(final_ridge_model, s = final_lambda_ridge))
+holdout_coefs <- as.matrix(ridge_test_coefs)
 
-# View model coefficients
-print(coef(ridge_model))
+# Remove the intercept term (row with name "(Intercept)") from both coefficient matrices
+training_coefs <- training_coefs[rownames(training_coefs) != "(Intercept)", , drop = FALSE]
+holdout_coefs <- holdout_coefs[rownames(holdout_coefs) != "(Intercept)", , drop = FALSE]
+
+# Ensure feature names match between the training and holdout coefficients
+common_features <- intersect(rownames(training_coefs), rownames(holdout_coefs))
+
+# Subset coefficients to only include common features
+training_coefs <- training_coefs[common_features, , drop = FALSE]
+holdout_coefs <- holdout_coefs[common_features, , drop = FALSE]
+
+# Combine coefficients into a comparison data frame
+comparison <- data.frame(
+  Predictor = common_features,
+  Training_Coefs = round(as.numeric(training_coefs), 8),
+  Holdout_Coefs = round(as.numeric(holdout_coefs), 8)
+)
+
+# Sort the comparison data frame in decreasing order of Training_Coefs
+comparison <- comparison[order(-comparison$Training_Coefs), ]
+
+# Set options to avoid scientific notation
+options(scipen = 999)
+
+# Display the comparison
+print("Comparison of coefficients from training and holdout data:")
+print(comparison)
 
 ```
+
+```{r 2d bootstrap Ridge}
+
+# Function to fit Ridge on bootstrap sample and return coefficients
+bootstrap_ridge <- function(data, indices, lambda) {
+  # Bootstrap sampling
+  boot_sample <- data[indices, ]
+
+  # Create the model matrix from the bootstrap sample
+  X_boot <- model.matrix(Log_Revenue_Total ~ ., data = boot_sample)[, -1]
+  y_boot <- boot_sample$Log_Revenue_Total
+
+  # Ensure that X_boot has the same column names as the original data
+  full_X <- model.matrix(Log_Revenue_Total ~ ., data = data)[, -1]
+  full_coefs <- numeric(ncol(full_X))  # Create a vector for all predictors (set to 0 initially)
+  names(full_coefs) <- colnames(full_X)  # Name it with the original predictor names
+
+  # Fit Ridge model
+  ridge_fit <- glmnet(X_boot, y_boot, alpha = 0, lambda = lambda, standardize = TRUE)
+
+  # Extract coefficients (excluding intercept)
+  coefs <- as.vector(coef(ridge_fit)[-1])  # Exclude intercept
+  coef_names <- colnames(X_boot)  # Names of predictors in this bootstrap sample
+
+  # Assign the coefficients to the correct positions
+  full_coefs[coef_names] <- coefs
+
+  return(full_coefs)
+}
+
+# Set up bootstrap
+set.seed(123)
+n_boot <- 1000  # Number of bootstrap samples
+X_train <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]
+y_train <- df_pred$Log_Revenue_Total
+
+# Perform bootstrap
+bootstrap_results <- boot(
+  data = df_pred,
+  statistic = bootstrap_ridge,
+  R = n_boot,
+  lambda = final_lambda_ridge
+)
+
+# Compute confidence intervals for each coefficient
+bootstrap_cis <- apply(bootstrap_results$t, 2, function(coefs) {
+  quantile(coefs, probs = c(0.025, 0.975))
+})
+
+# Format as a data frame
+cis_df <- data.frame(
+  Predictor = colnames(X_train),
+  CI_Lower = bootstrap_cis[1, ],
+  CI_Upper = bootstrap_cis[2, ]
+)
+
+## Compare to the confidence intervals computed in the standard regression output
+
+# Identify numeric predictors
+numeric_predictors <- sapply(df_pred[predictor_columns], is.numeric)
+
+# Standardize only numeric predictors
+df_pred_standardized <- df_pred
+df_pred_standardized[names(numeric_predictors)[numeric_predictors]] <- scale(
+  df_pred[names(numeric_predictors)[numeric_predictors]]
+)
+
+# # See that non-numeric columns remain unchanged
+# print("Standardized dataset:")
+# print(head(df_pred_standardized))
+
+# Fit OLS regression with standardized predictors
+ols_model_standardized <- lm(Log_Revenue_Total ~ ., data = df_pred_standardized)
+
+# Extract OLS confidence intervals
+ols_cis_standardized <- confint(ols_model_standardized, level = 0.95)
+ols_cis_df_standardized <- data.frame(
+  Predictor = rownames(ols_cis_standardized),
+  OLS_CI_Lower_Standardized = ols_cis_standardized[, 1],
+  OLS_CI_Upper_Standardized = ols_cis_standardized[, 2]
+)
+
+# Merge with Bootstrap confidence intervals for comparison
+comparison_ridge_OLS <- merge(
+  cis_df,  # Bootstrap confidence intervals from Ridge
+  ols_cis_df_standardized,
+  by = "Predictor",
+  all.x = TRUE
+)
+
+# Display the comparison
+print("Comparison of Bootstrap and Standardized OLS Confidence Intervals:")
+print(comparison_ridge_OLS)
+
+# visualize the difference in Ridge vs. OLS CIs
+
+# Calculate confidence interval sizes for Ridge and OLS
+comparison_ridge_OLS$Ridge_CI_Size <- comparison_ridge_OLS$CI_Upper - comparison_ridge_OLS$CI_Lower
+comparison_ridge_OLS$OLS_CI_Size <- comparison_ridge_OLS$OLS_CI_Upper_Standardized - comparison_ridge_OLS$OLS_CI_Lower_Standardized
+
+# Reshape the data for plotting
+ci_data_ridge <- comparison_ridge_OLS %>%
+  select(Predictor, Ridge_CI_Size, OLS_CI_Size) %>%
+  pivot_longer(cols = c(Ridge_CI_Size, OLS_CI_Size), 
+               names_to = "Method", 
+               values_to = "CI_Size")
+
+# Rename methods for better readability
+ci_data_ridge$Method <- ifelse(ci_data_ridge$Method == "Ridge_CI_Size", "Ridge", "OLS")
+
+# Create the density plot
+ggplot(ci_data_ridge, aes(x = CI_Size, fill = Method)) +
+  geom_density(alpha = 0.5) +
+  scale_x_log10() +
+  labs(
+    title = "Density Plot of Confidence Interval Sizes: Ridge vs. OLS",
+    x = "Confidence Interval Size",
+    y = "Density"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(size = 14, face = "bold"),
+    axis.title = element_text(size = 12),
+    legend.title = element_blank()
+  )
+
+# violin plot
+ggplot(ci_data_ridge, aes(x = Method, y = CI_Size)) +
+  geom_violin(fill = "lightblue") +
+  geom_boxplot(width = 0.1, fill = "white") +
+  theme_minimal() +
+  labs(title = "Distribution of CI_Size by Group", x = "Group", y = "CI_Size")
+
+```
+
+```{r 2c Fit to test data Lasso}
+
+# Define the columns to include in df_pred
+column_names <- c(
+  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Sources.Purchased",
+  "Sources.Other", "Uses.Retail", "Uses.Resale", "Uses.No.Charge", "Uses.Consumed", "Uses.Losses",
+  "Retail.Residential.Customers", "Retail.Commercial.Customers", "Retail.Industrial.Customers",
+  "Retail.Transportation.Customers", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type"
+)
+
+# save a copy of the holdout dataset to be transformed/cleaned 
+# Create df_test with only the specified columns from holdout
+df_test <- holdout[, column_names]
+
+# Log-transform the Revenue.Total column in the test set to match training
+df_test$Log_Revenue_Total <- log1p(df_test$Revenue.Total)
+
+# # Extract predictors (X) and response variable (y) for the test set
+# X_test <- model.matrix(Log_Revenue_Total ~ ., data = df_test)[, -1]
+# y_test <- df_test$Log_Revenue_Total
+
+# Refit the Lasso model on the test set using the final lambda from training
+lasso_test <- glmnet(X_test, y_test, alpha = 1, lambda = final_lambda_lasso, standardize = TRUE)
+
+# Extract coefficients from the test set
+lasso_test_coefs <- coef(lasso_test)
+
+# Extract coefficients from training and holdout models
+training_coefs <- as.matrix(coef(lasso_model$finalModel, s = final_lambda_lasso))
+holdout_coefs <- as.matrix(lasso_test_coefs)
+
+# Remove the intercept term (row with name "(Intercept)") from both coefficient matrices
+training_coefs <- training_coefs[rownames(training_coefs) != "(Intercept)", , drop = FALSE]
+holdout_coefs <- holdout_coefs[rownames(holdout_coefs) != "(Intercept)", , drop = FALSE]
+
+# Ensure feature names match between the training and holdout coefficients
+common_features <- intersect(rownames(training_coefs), rownames(holdout_coefs))
+
+# Subset coefficients to only include common features
+training_coefs <- training_coefs[common_features, , drop = FALSE]
+holdout_coefs <- holdout_coefs[common_features, , drop = FALSE]
+
+# Combine coefficients into a comparison data frame
+comparison <- data.frame(
+  Predictor = common_features,
+  Training_Coefs = round(as.numeric(training_coefs), 8),
+  Holdout_Coefs = round(as.numeric(holdout_coefs), 8)
+)
+
+# Sort the comparison data frame in decreasing order of Training_Coefs
+comparison <- comparison[order(-comparison$Training_Coefs), ]
+
+# Set options to avoid scientific notation
+options(scipen = 999)
+
+# Display the comparison
+print("Comparison of coefficients from training and holdout data:")
+print(comparison)
+
+```
+
+```{r 2d bootstrap Lasso}
+
+# Function to fit Lasso on bootstrap sample and return coefficients
+bootstrap_lasso <- function(data, indices, lambda) {
+  # Bootstrap sampling
+  boot_sample <- data[indices, ]
+
+  # Create the model matrix from the bootstrap sample
+  X_boot <- model.matrix(Log_Revenue_Total ~ ., data = boot_sample)[, -1]
+  y_boot <- boot_sample$Log_Revenue_Total
+
+  # Ensure that X_boot has the same column names as the original data
+  full_X <- model.matrix(Log_Revenue_Total ~ ., data = data)[, -1]
+  full_coefs <- numeric(ncol(full_X))  # Create a vector for all predictors (set to 0 initially)
+  names(full_coefs) <- colnames(full_X)  # Name it with the original predictor names
+
+  # Fit Lasso model
+  lasso_fit <- glmnet(X_boot, y_boot, alpha = 1, lambda = lambda, standardize = TRUE)
+
+  # Extract coefficients (excluding intercept)
+  coefs <- as.vector(coef(lasso_fit)[-1])  # Exclude intercept
+  coef_names <- colnames(X_boot)  # Names of predictors in this bootstrap sample
+
+  # Assign the coefficients to the correct positions
+  full_coefs[coef_names] <- coefs
+
+  return(full_coefs)
+}
+
+# Set up bootstrap
+set.seed(123)
+n_boot <- 1000  # Number of bootstrap samples
+X_train <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]
+y_train <- df_pred$Log_Revenue_Total
+
+# Perform bootstrap
+bootstrap_results <- boot(
+  data = df_pred,
+  statistic = bootstrap_lasso,
+  R = n_boot,
+  lambda = final_lambda_lasso
+)
+
+# Compute confidence intervals for each coefficient
+bootstrap_cis <- apply(bootstrap_results$t, 2, function(coefs) {
+  quantile(coefs, probs = c(0.025, 0.975))
+})
+
+# Format as a data frame
+cis_df <- data.frame(
+  Predictor = colnames(X_train),
+  CI_Lower = bootstrap_cis[1, ],
+  CI_Upper = bootstrap_cis[2, ]
+)
+
+## Compare to the confidence intervals computed in the standard regression output
+
+# Identify numeric predictors
+numeric_predictors <- sapply(df_pred[predictor_columns], is.numeric)
+
+# Standardize only numeric predictors
+df_pred_standardized <- df_pred
+df_pred_standardized[names(numeric_predictors)[numeric_predictors]] <- scale(
+  df_pred[names(numeric_predictors)[numeric_predictors]]
+)
+
+# # See that non-numeric columns remain unchanged
+# print("Standardized dataset:")
+# print(head(df_pred_standardized))
+
+# Fit OLS regression with standardized predictors
+ols_model_standardized <- lm(Log_Revenue_Total ~ ., data = df_pred_standardized)
+
+# Extract OLS confidence intervals
+ols_cis_standardized <- confint(ols_model_standardized, level = 0.95)
+ols_cis_df_standardized <- data.frame(
+  Predictor = rownames(ols_cis_standardized),
+  OLS_CI_Lower_Standardized = ols_cis_standardized[, 1],
+  OLS_CI_Upper_Standardized = ols_cis_standardized[, 2]
+)
+
+# Merge with Bootstrap confidence intervals for comparison
+comparison_lasso_OLS <- merge(
+  cis_df,  # Bootstrap confidence intervals from Lasso
+  ols_cis_df_standardized,
+  by = "Predictor",
+  all.x = TRUE
+)
+
+# Display the comparison
+print("Comparison of Bootstrap and Standardized OLS Confidence Intervals:")
+print(comparison_lasso_OLS)
+
+# visualize the difference in Lasso vs. OLS CIs
+
+# Calculate confidence interval sizes for Lasso and OLS
+comparison_lasso_OLS$Lasso_CI_Size <- comparison_lasso_OLS$CI_Upper - comparison_lasso_OLS$CI_Lower
+comparison_lasso_OLS$OLS_CI_Size <- comparison_lasso_OLS$OLS_CI_Upper_Standardized - comparison_lasso_OLS$OLS_CI_Lower_Standardized
+
+# Reshape the data for plotting
+ci_data_lasso <- comparison_lasso_OLS %>%
+  select(Predictor, Lasso_CI_Size, OLS_CI_Size) %>%
+  pivot_longer(cols = c(Lasso_CI_Size, OLS_CI_Size), 
+               names_to = "Method", 
+               values_to = "CI_Size")
+
+# Rename methods for better readability
+ci_data_lasso$Method <- ifelse(ci_data_lasso$Method == "Lasso_CI_Size", "Lasso", "OLS")
+
+# Create the density plot
+ggplot(ci_data_lasso, aes(x = CI_Size, fill = Method)) +
+  geom_density(alpha = 0.5) +
+  scale_x_log10() +
+  labs(
+    title = "Density Plot of Confidence Interval Sizes: Lasso vs. OLS",
+    x = "Confidence Interval Size",
+    y = "Density"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(size = 14, face = "bold"),
+    axis.title = element_text(size = 12),
+    legend.title = element_blank()
+  )
+
+# violin plot
+ggplot(ci_data_lasso, aes(x = Method, y = CI_Size)) +
+  geom_violin(fill = "lightblue") +
+  geom_boxplot(width = 0.1, fill = "white") +
+  theme_minimal() +
+  labs(title = "Distribution of CI_Size by Group", x = "Group", y = "CI_Size")
+
+```

### 2024-12-02 MS&E 226_Project Part 1_Code.Rmd
code with ridge regression as model selected

@@
 View(region_means)
-variables_to_plot <- names(state_means)[-1]  # Exclude Utility.State
-plot_list <- list()
 ```
@@
 ```{r Utility.State plots, fig.show='hide', results='hide'}
+variables_to_plot <- names(state_means)[-1]  # Exclude Utility.State
+plot_list <- list()
 # Loop through each variable and create a plot

### 2024-12-02 MS&E 226_Project Part 2_Code.Rmd
code with ridge regression as model selected

@@
 
+### Regression
+
+```{r Regression prediction with winning modeling strategy, Random Forest Model}
+
+column_names <- c(
+  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Sources.Purchased",
+  "Sources.Other", "Uses.Retail", "Uses.Resale", "Uses.No.Charge", "Uses.Consumed", "Uses.Losses",
+  "Retail.Residential.Customers", "Retail.Commercial.Customers", "Retail.Industrial.Customers",
+  "Retail.Transportation.Customers", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type"
+)
+
+# Select the relevant columns in the test set
+holdout_pred <- holdout[, column_names]
+# Apply the same transformation to the test set
+holdout_pred$Log_Revenue_Total <- log1p(holdout$Revenue.Total)
+
+# Predict on the test set
+rf_test_predictions <- predict(rf_model, newdata = holdout_pred)
+
+# Check for NA values in predictions or Log_Revenue_Total
+if (any(is.na(rf_test_predictions)) || any(is.na(holdout_pred$Log_Revenue_Total))) {
+    stop("Predictions or Log_Revenue_Total contain NA values.")
+}
+
+# Calculate RMSE
+rf_test_rmse <- sqrt(mean((rf_test_predictions - holdout_pred$Log_Revenue_Total)^2))
+cat(paste("Random Forest Test RMSE:", rf_test_rmse, "\n"))
+
+```
+
 ### Classification
@@
 ## 2: Inference
+
+#### (a)
+```{r 2a}
+# Extract the final Ridge model using the best lambda
+final_ridge_model <- ridge_model$finalModel
+
+# Extract coefficients from the final model
+coefficients <- coef(final_ridge_model, s = final_lambda_ridge)
+
+# Print the coefficients
+print(coefficients)
+
+```
+
+#### (b)
+```{r}
+# Select a subset of covariates
+subset_column_names <- c(  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type")
+
+# Create a new data frame with the subset of columns
+df_subset <- df[, subset_column_names]
+
+df_subset$Log_Revenue_Total <- log1p(df_subset$Revenue.Total)
+
+# Train the Ridge Regression model using the subset of covariates
+ridge_subset_model <- train(Log_Revenue_Total ~ ., 
+                            data = df_subset, 
+                            method = "glmnet",  
+                            trControl = train_control,
+                            metric = "RMSE", 
+                            preProcess = "scale",  
+                            tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 1, length = 100)))
+
+```
+```{r}
+# Get the coefficients for the best model at the selected lambda
+best_lambda <- ridge_model$bestTune$lambda
+best_model_coeffs <- coef(ridge_model$finalModel, s = best_lambda)
+
+# Print all coefficients for the best Ridge Regression model
+cat("All coefficients for the best Ridge Regression model:\n")
+print(best_model_coeffs)
+
+# Get coefficients for the subset model
+best_lambda_subset <- ridge_subset_model$bestTune$lambda
+best_model_coeffs_subset <- coef(ridge_subset_model$finalModel, s = best_lambda_subset)
+
+# Print all coefficients for the subset model
+cat("All coefficients for the Ridge Regression model with a subset of covariates:\n")
+print(best_model_coeffs_subset)
+```
 

### 2024-12-02 MS&E 226_Project Part 2_Code.Rmd
comments

@@
 ```
-
-# 2: Data Pre-processing
+#Projetc Part 1
+
+## 2. Data Pre-processing
 
@@
 
-# 3: Hold Out Set
+## 3. Hold Out Set
 
@@
 
-# 5: Data Exploration
-
-## Correlations
+## 5. Data Exploration
+
+### Correlations
 
@@
 
-## Subgroups in data
+### Subgroups in data
 
@@
 
-# 6: Regression
-
-## Data Transformations
+## 6. Regression
+
+### Data Transformations
 
@@
 
-## Baseline Model
+### Baseline Model
 
@@
 
-## Cross Validation setup
+### Cross Validation setup
 
@@
 
-## Model 1: Ridge Regression
+### Model 1: Ridge Regression
 
@@
 
-## Model 2: Lasso Regression
+### Model 2: Lasso Regression
 
@@
 
-## Model 3: Random Forest
+### Model 3: Random Forest
 
@@
 ```
-
-# 7: Classification
-
-## Data Cleaning
+## 7. Classification
+
+### Data Cleaning
 
@@
 
-## Classifier 1: K-NN
+### Classifier 1: K-NN
 
@@
 
-## Classifier 2: Logistic Regression with Feature Selection
+### Classifier 2: Logistic Regression with Feature Selection
 
@@
 
-## Classifier 3: Decision Tree
+### Classifier 3: Decision Tree
 
@@
 
-## Comparing Classifiers
+### Comparing Classifiers
 
@@
 
-## 1: Prediction on the holdout set
+## 1. Prediction on the holdout set
 
@@
 ### Classification
-
-#### a)
 
@@
 
-## 2: Inference
-
-#### (a)
+## 2. Inference
+
+### Statistically significant Coeficients
 ```{r 2a}
@@
 
-#### (b)
+### Subset of covariates
 ```{r}
@@
 ```
+
 ```{r}
@@
 
-#### (c)
+### Fit on Holdout set
 
@@
 
+### Bootstrap
 ```{r 2d bootstrap Ridge}

### 2024-12-02 MS&E 226_Project Part 2_Code.Rmd
OLS on Lasso covariates

--- 2024-12-02 MS&E 226_Project Part 2_Code.Rmd
@@
 
+```{r create OLS model using covariates selectedin Lasso}
+
+# Generate the final Lasso model
+
+# Extract predictors (X) and response variable (y) for the train set
+X_train <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]
+y_train <- df_pred$Log_Revenue_Total
+
+# Refit the Lasso model on the test set using the final lambda from training
+final_lasso_model <- glmnet(X_train, y_train, alpha = 1, lambda = final_lambda_lasso, standardize = TRUE)
+
+# Extract non-zero coefficients from the final Lasso model
+lasso_coefficients <- coef(final_lasso_model)
+lasso_coefficients <- as.matrix(lasso_coefficients)  # Convert to a matrix
+nonzero_covariates <- rownames(lasso_coefficients)[lasso_coefficients != 0]
+nonzero_covariates <- nonzero_covariates[nonzero_covariates != "(Intercept)"] # Exclude intercept
+
+# Identify all one-hot encoded categorical variables in this list and remove all but one
+remove_after_capital <- function(x) {
+  # Use a regular expression to find a lowercase letter followed by an uppercase letter
+  gsub("([a-z])[A-Z].*", "\\1", x)
+}
+
+# Apply the function to the vector of strings
+unique_nonzero_covariates <- sapply(nonzero_covariates, remove_after_capital)
+
+# Print the modified strings
+print(unique_nonzero_covariates)
+
+# Identify unique strings (appear exactly once)
+unique_nonzero_covariates <- names(table(unique_nonzero_covariates)[table(unique_nonzero_covariates) == 1])
+
+# Print the unique strings
+print(unique_nonzero_covariates)
+
+# Subset the dataframe with the selected covariates
+df_selected <- df_pred[, c("Log_Revenue_Total", unique_nonzero_covariates)]
+
+# Create the formula for OLS regression
+ols_formula <- as.formula(paste("Log_Revenue_Total ~", paste(unique_nonzero_covariates, collapse = " + ")))
+
+# Run the OLS regression
+ols_train <- lm(ols_formula, data = df_selected)
+
+# Print summary of the OLS regression model
+summary(ols_train)
+
+```
+
 #### (a)
@@
 #### (c)
-
-```{r 2c Fit to test data Ridge}
-
-# manually refit final ridge model
-
-# Refit the model manually using glmnet
-X_train <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]
-y_train <- df_pred$Log_Revenue_Total
-
-final_ridge_model <- glmnet(X_train, y_train, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
+```{r 2c Fit to test data OLS}
 
@@
 
-# Extract predictors (X) and response variable (y) for the test set
-X_test <- model.matrix(Log_Revenue_Total ~ ., data = df_test)[, -1]
-y_test <- df_test$Log_Revenue_Total
-
-# Refit the ridge model on the test set using the final lambda from training
-ridge_test <- glmnet(X_test, y_test, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
+# Refit the OLS model on the test set 
+ols_test <- lm(ols_formula, data = df_test)
 
 # Extract coefficients from the test set
-ridge_test_coefs <- coef(ridge_test)
+ols_test_coefs <- coef(ols_test)
 
 # Extract coefficients from training and holdout models
-training_coefs <- as.matrix(coef(final_ridge_model, s = final_lambda_ridge))
-holdout_coefs <- as.matrix(ridge_test_coefs)
+training_coefs <- as.matrix(coef(ols_train))
+holdout_coefs <- as.matrix(ols_test_coefs)
 
@@
 
-```{r 2d bootstrap Ridge}
+```{r 2d bootstrap OLS}
 
@@
 # Identify numeric predictors
-numeric_predictors <- sapply(df_pred[predictor_columns], is.numeric)
+numeric_predictors <- sapply(df_pred[column_names], is.numeric)
 
@@
 
-```{r 2c Fit to test data Lasso}
+```{r 2c Fit to test data Ridge}
+
+# manually refit final ridge model
+
+# Refit the model manually using glmnet
+X_train <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]
+y_train <- df_pred$Log_Revenue_Total
+
+final_ridge_model <- glmnet(X_train, y_train, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
 
@@
 
+# Extract predictors (X) and response variable (y) for the test set
+X_test <- model.matrix(Log_Revenue_Total ~ ., data = df_test)[, -1]
+y_test <- df_test$Log_Revenue_Total
+
+# Refit the ridge model on the test set using the final lambda from training
+ridge_test <- glmnet(X_test, y_test, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
+
+# Extract coefficients from the test set
+ridge_test_coefs <- coef(ridge_test)
+
+# Extract coefficients from training and holdout models
+training_coefs <- as.matrix(coef(final_ridge_model, s = final_lambda_ridge))
+holdout_coefs <- as.matrix(ridge_test_coefs)
+
+# Remove the intercept term (row with name "(Intercept)") from both coefficient matrices
+training_coefs <- training_coefs[rownames(training_coefs) != "(Intercept)", , drop = FALSE]
+holdout_coefs <- holdout_coefs[rownames(holdout_coefs) != "(Intercept)", , drop = FALSE]
+
+# Ensure feature names match between the training and holdout coefficients
+common_features <- intersect(rownames(training_coefs), rownames(holdout_coefs))
+
+# Subset coefficients to only include common features
+training_coefs <- training_coefs[common_features, , drop = FALSE]
+holdout_coefs <- holdout_coefs[common_features, , drop = FALSE]
+
+# Combine coefficients into a comparison data frame
+comparison <- data.frame(
+  Predictor = common_features,
+  Training_Coefs = round(as.numeric(training_coefs), 8),
+  Holdout_Coefs = round(as.numeric(holdout_coefs), 8)
+)
+
+# Sort the comparison data frame in decreasing order of Training_Coefs
+comparison <- comparison[order(-comparison$Training_Coefs), ]
+
+# Set options to avoid scientific notation
+options(scipen = 999)
+
+# Display the comparison
+print("Comparison of coefficients from training and holdout data:")
+print(comparison)
+
+```
+
+```{r 2d bootstrap Ridge}
+
+# Function to fit Ridge on bootstrap sample and return coefficients
+bootstrap_ridge <- function(data, indices, lambda) {
+  # Bootstrap sampling
+  boot_sample <- data[indices, ]
+
+  # Create the model matrix from the bootstrap sample
+  X_boot <- model.matrix(Log_Revenue_Total ~ ., data = boot_sample)[, -1]
+  y_boot <- boot_sample$Log_Revenue_Total
+
+  # Ensure that X_boot has the same column names as the original data
+  full_X <- model.matrix(Log_Revenue_Total ~ ., data = data)[, -1]
+  full_coefs <- numeric(ncol(full_X))  # Create a vector for all predictors (set to 0 initially)
+  names(full_coefs) <- colnames(full_X)  # Name it with the original predictor names
+
+  # Fit Ridge model
+  ridge_fit <- glmnet(X_boot, y_boot, alpha = 0, lambda = lambda, standardize = TRUE)
+
+  # Extract coefficients (excluding intercept)
+  coefs <- as.vector(coef(ridge_fit)[-1])  # Exclude intercept
+  coef_names <- colnames(X_boot)  # Names of predictors in this bootstrap sample
+
+  # Assign the coefficients to the correct positions
+  full_coefs[coef_names] <- coefs
+
+  return(full_coefs)
+}
+
+# Set up bootstrap
+set.seed(123)
+n_boot <- 1000  # Number of bootstrap samples
+X_train <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]
+y_train <- df_pred$Log_Revenue_Total
+
+# Perform bootstrap
+bootstrap_results <- boot(
+  data = df_pred,
+  statistic = bootstrap_ridge,
+  R = n_boot,
+  lambda = final_lambda_ridge
+)
+
+# Compute confidence intervals for each coefficient
+bootstrap_cis <- apply(bootstrap_results$t, 2, function(coefs) {
+  quantile(coefs, probs = c(0.025, 0.975))
+})
+
+# Format as a data frame
+cis_df <- data.frame(
+  Predictor = colnames(X_train),
+  CI_Lower = bootstrap_cis[1, ],
+  CI_Upper = bootstrap_cis[2, ]
+)
+
+## Compare to the confidence intervals computed in the standard regression output
+
+# Identify numeric predictors
+numeric_predictors <- sapply(df_pred[column_names], is.numeric)
+
+# Standardize only numeric predictors
+df_pred_standardized <- df_pred
+df_pred_standardized[names(numeric_predictors)[numeric_predictors]] <- scale(
+  df_pred[names(numeric_predictors)[numeric_predictors]]
+)
+
+# # See that non-numeric columns remain unchanged
+# print("Standardized dataset:")
+# print(head(df_pred_standardized))
+
+# Fit OLS regression with standardized predictors
+ols_model_standardized <- lm(Log_Revenue_Total ~ ., data = df_pred_standardized)
+
+# Extract OLS confidence intervals
+ols_cis_standardized <- confint(ols_model_standardized, level = 0.95)
+ols_cis_df_standardized <- data.frame(
+  Predictor = rownames(ols_cis_standardized),
+  OLS_CI_Lower_Standardized = ols_cis_standardized[, 1],
+  OLS_CI_Upper_Standardized = ols_cis_standardized[, 2]
+)
+
+# Merge with Bootstrap confidence intervals for comparison
+comparison_ridge_OLS <- merge(
+  cis_df,  # Bootstrap confidence intervals from Ridge
+  ols_cis_df_standardized,
+  by = "Predictor",
+  all.x = TRUE
+)
+
+# Display the comparison
+print("Comparison of Bootstrap and Standardized OLS Confidence Intervals:")
+print(comparison_ridge_OLS)
+
+# visualize the difference in Ridge vs. OLS CIs
+
+# Calculate confidence interval sizes for Ridge and OLS
+comparison_ridge_OLS$Ridge_CI_Size <- comparison_ridge_OLS$CI_Upper - comparison_ridge_OLS$CI_Lower
+comparison_ridge_OLS$OLS_CI_Size <- comparison_ridge_OLS$OLS_CI_Upper_Standardized - comparison_ridge_OLS$OLS_CI_Lower_Standardized
+
+# Reshape the data for plotting
+ci_data_ridge <- comparison_ridge_OLS %>%
+  select(Predictor, Ridge_CI_Size, OLS_CI_Size) %>%
+  pivot_longer(cols = c(Ridge_CI_Size, OLS_CI_Size), 
+               names_to = "Method", 
+               values_to = "CI_Size")
+
+# Rename methods for better readability
+ci_data_ridge$Method <- ifelse(ci_data_ridge$Method == "Ridge_CI_Size", "Ridge", "OLS")
+
+# Create the density plot
+ggplot(ci_data_ridge, aes(x = CI_Size, fill = Method)) +
+  geom_density(alpha = 0.5) +
+  scale_x_log10() +
+  labs(
+    title = "Density Plot of Confidence Interval Sizes: Ridge vs. OLS",
+    x = "Confidence Interval Size",
+    y = "Density"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(size = 14, face = "bold"),
+    axis.title = element_text(size = 12),
+    legend.title = element_blank()
+  )
+
+# violin plot
+ggplot(ci_data_ridge, aes(x = Method, y = CI_Size)) +
+  geom_violin(fill = "lightblue") +
+  geom_boxplot(width = 0.1, fill = "white") +
+  theme_minimal() +
+  labs(title = "Distribution of CI_Size by Group", x = "Group", y = "CI_Size")
+
+```
+
+```{r 2c Fit to test data Lasso}
+
+# Define the columns to include in df_pred
+column_names <- c(
+  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Sources.Purchased",
+  "Sources.Other", "Uses.Retail", "Uses.Resale", "Uses.No.Charge", "Uses.Consumed", "Uses.Losses",
+  "Retail.Residential.Customers", "Retail.Commercial.Customers", "Retail.Industrial.Customers",
+  "Retail.Transportation.Customers", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type"
+)
+
+# save a copy of the holdout dataset to be transformed/cleaned 
+# Create df_test with only the specified columns from holdout
+df_test <- holdout[, column_names]
+
+# Log-transform the Revenue.Total column in the test set to match training
+df_test$Log_Revenue_Total <- log1p(df_test$Revenue.Total)
+
 # # Extract predictors (X) and response variable (y) for the test set

### 2024-12-03 MS&E 226_Project Part 2_Code.Rmd
ols model as chosen model

@@
 
-# 2: Data Pre-processing
+#Project Part 1
+
+## 2. Data Pre-processing
 
@@
 
-# 7: Classification
-
-## Data Cleaning
+## 7. Classification
+
+### Data Cleaning
 
@@
 
-## 2: Inference
+## 2. Inference
 
@@
 # Print summary of the OLS regression model
-summary(ols_train)
-
-```
-
-#### (a)
+ols_summary <- summary(ols_train)
+
+```
+
+### Statistically Significant Coefficients
 ```{r 2a}
-# Extract the final Ridge model using the best lambda
-final_ridge_model <- ridge_model$finalModel
-
-# Extract coefficients from the final model
-coefficients <- coef(final_ridge_model, s = final_lambda_ridge)
-
-# Print the coefficients
+coefficients <- ols_summary$coefficients
 print(coefficients)
 
-```
-
-#### (b)
+# Extract and identify statistically significant coefficients (p-value < 0.05)
+significant_coeffs <- coefficients[coefficients[, 4] < 0.05, ]
+
+# Print the statistically significant coefficients and their p-values
+print("Statistically significant coefficients:")
+print(significant_coeffs)
+
+# Now, let's extract the names of the statistically significant variables
+significant_variable_names <- rownames(significant_coeffs)
+print("Significant variables:")
+print(significant_variable_names)
+
+```
+
+### Subset of covariates model
+
 ```{r}
-# Select a subset of covariates
-subset_column_names <- c(  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type")
+# Select a subset of covariates (as you did before)
+subset_column_names <- c("Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type")
 
@@
 
+# Create the log-transformed revenue variable
 df_subset$Log_Revenue_Total <- log1p(df_subset$Revenue.Total)
-
-# Train the Ridge Regression model using the subset of covariates
-ridge_subset_model <- train(Log_Revenue_Total ~ ., 
-                            data = df_subset, 
-                            method = "glmnet",  
-                            trControl = train_control,
-                            metric = "RMSE", 
-                            preProcess = "scale",  
-                            tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 1, length = 100)))
-
-```
+df_subset$Revenue.Total <- NULL
+
+# Create the formula for OLS regression
+ols_formula <- as.formula(paste("Log_Revenue_Total ~", paste(subset_column_names[-6], collapse = " + ")))
+
+# Train the OLS regression model
+ols_model <- lm(ols_formula, data = df_subset)
+
+# Print the summary of the OLS model
+cat("OLS model coefficients summary:\n")
+summary(ols_model)
+
+# Extract and print the coefficients
+cat("OLS model coefficients:\n")
+print(coef(ols_model))
+
+# Optional: Check the significance of each coefficient using the summary output
+cat("OLS model coefficient significance:\n")
+summary(ols_model)$coefficients[, 4]  # p-values of the coefficients
+
+```
+
 ```{r}
-# Get the coefficients for the best model at the selected lambda
-best_lambda <- ridge_model$bestTune$lambda
-best_model_coeffs <- coef(ridge_model$finalModel, s = best_lambda)
-
-# Print all coefficients for the best Ridge Regression model
-cat("All coefficients for the best Ridge Regression model:\n")
-print(best_model_coeffs)
-
-# Get coefficients for the subset model
-best_lambda_subset <- ridge_subset_model$bestTune$lambda
-best_model_coeffs_subset <- coef(ridge_subset_model$finalModel, s = best_lambda_subset)
-
-# Print all coefficients for the subset model
-cat("All coefficients for the Ridge Regression model with a subset of covariates:\n")
-print(best_model_coeffs_subset)
+# Get the summary of the OLS model
+ols_summary <- summary(ols_model)
+
+# Extract the p-values of the coefficients
+p_values <- ols_summary$coefficients[, 4]
+
+# Set a significance threshold (e.g., 0.05)
+significance_threshold <- 0.05
+
+# Identify significant coefficients based on p-value
+significant_coeffs <- names(p_values)[p_values < significance_threshold]
+
+# Print the significant coefficients
+cat("Significant coefficients based on p-value < 0.05:\n")
+print(significant_coeffs)
+
 ```

### 2024-12-03 MS&E 226_Project Part 2_Code.Rmd
2c and 2d with OLS fit from lasso coefficients

--- 2024-12-02 MS&E 226_Project Part 2_Code.Rmd
@@
 #### (c)
+
 ```{r 2c Fit to test data OLS}
@@
 
-# save a copy of the holdout dataset to be transformed/cleaned 
+# Save a copy of the holdout dataset to be transformed/cleaned 
 # Create df_test with only the specified columns from holdout
@@
 
-# Extract coefficients from the test set
-ols_test_coefs <- coef(ols_test)
-
-# Extract coefficients from training and holdout models
-training_coefs <- as.matrix(coef(ols_train))
-holdout_coefs <- as.matrix(ols_test_coefs)
-
-# Remove the intercept term (row with name "(Intercept)") from both coefficient matrices
-training_coefs <- training_coefs[rownames(training_coefs) != "(Intercept)", , drop = FALSE]
-holdout_coefs <- holdout_coefs[rownames(holdout_coefs) != "(Intercept)", , drop = FALSE]
-
-# Ensure feature names match between the training and holdout coefficients
-common_features <- intersect(rownames(training_coefs), rownames(holdout_coefs))
+# Extract p-values from the training and test models
+training_pvalues <- summary(ols_train)$coefficients[, "Pr(>|t|)"]
+test_pvalues <- summary(ols_test)$coefficients[, "Pr(>|t|)"]
+
+# Remove the intercept term from p-values
+training_pvalues <- training_pvalues[names(training_pvalues) != "(Intercept)"]
+test_pvalues <- test_pvalues[names(test_pvalues) != "(Intercept)"]
+
+# Ensure feature names match between the training and holdout p-values
+common_features <- intersect(names(training_pvalues), names(test_pvalues))
+
+# Subset p-values to only include common features
+training_pvalues <- training_pvalues[common_features]
+test_pvalues <- test_pvalues[common_features]
+
+# Combine p-values into a comparison data frame
+comparison <- data.frame(
+  Predictor = common_features,
+  Training_Pvalues = round(training_pvalues, 8),
+  Test_Pvalues = round(test_pvalues, 8)
+)
+
+# Sort the comparison data frame in increasing order of Training_Pvalues
+comparison <- comparison[order(comparison$Training_Pvalues), ]
+
+# Set options to avoid scientific notation
+options(scipen = 999)
+
+# Display the comparison
+print("Comparison of p-values from training and test data:")
+print(comparison)
+
+## compare coefficients
 
@@
 
-# Set options to avoid scientific notation
-options(scipen = 999)
-
 # Display the comparison
@@
 ```{r 2d bootstrap OLS}
+
+# Function to fit OLS on bootstrap sample and return coefficients
+bootstrap_ols <- function(data, indices) {
+  # Bootstrap sampling
+  boot_sample <- data[indices, ]
+
+  # Fit OLS model
+  ols_fit <- lm(ols_formula, data = boot_sample)
+
+  # Extract coefficients
+  coefs <- coef(ols_fit)
+
+  # Return coefficients (excluding intercept if not needed)
+  return(coefs)
+}
+
+# Set up bootstrap
+set.seed(123)
+n_boot <- 1000  # Number of bootstrap samples
+
+# Perform bootstrap
+bootstrap_results <- boot(
+  data = df_pred,
+  statistic = bootstrap_ols,
+  R = n_boot
+)
+
+# Compute confidence intervals for each coefficient
+bootstrap_cis <- apply(bootstrap_results$t, 2, function(coefs) {
+  quantile(coefs, probs = c(0.025, 0.975))
+})
+
+# Format as a data frame
+cis_df <- data.frame(
+  Predictor = names(bootstrap_results$t0),
+  CI_Lower = bootstrap_cis[1, ],
+  CI_Upper = bootstrap_cis[2, ]
+)
+
+# Extract OLS confidence intervals from the standard regression output
+ols_cis_standardized <- confint(ols_test, level = 0.95)
+ols_cis_df_standardized <- data.frame(
+  Predictor = rownames(ols_cis_standardized),
+  OLS_CI_Lower_Standardized = ols_cis_standardized[, 1],
+  OLS_CI_Upper_Standardized = ols_cis_standardized[, 2]
+)
+
+# Merge with Bootstrap confidence intervals for comparison
+comparison_ols <- merge(
+  cis_df,  # Bootstrap confidence intervals from OLS
+  ols_cis_df_standardized,
+  by = "Predictor",
+  all.x = TRUE
+)
+
+# Display the comparison
+print("Comparison of Bootstrap and Standardized OLS Confidence Intervals:")
+print(comparison_ols)
+
+# Visualize the difference in Bootstrap vs. Standardized OLS CIs
+
+# Calculate confidence interval sizes for Bootstrap and OLS
+comparison_ols$Bootstrap_CI_Size <- comparison_ols$CI_Upper - comparison_ols$CI_Lower
+comparison_ols$OLS_CI_Size <- comparison_ols$OLS_CI_Upper_Standardized - comparison_ols$OLS_CI_Lower_Standardized
+
+# Reshape the data for plotting
+ci_data_ols <- comparison_ols %>%
+  select(Predictor, Bootstrap_CI_Size, OLS_CI_Size) %>%
+  pivot_longer(cols = c(Bootstrap_CI_Size, OLS_CI_Size), 
+               names_to = "Method", 
+               values_to = "CI_Size")
+
+# Rename methods for better readability
+ci_data_ols$Method <- ifelse(ci_data_ols$Method == "Bootstrap_CI_Size", "Bootstrap OLS", "Standard OLS Output")
+
+# Create the density plot
+ggplot(ci_data_ols, aes(x = CI_Size, fill = Method)) +
+  geom_density(alpha = 0.5) +
+  scale_x_log10() +
+  labs(
+    title = "Density Plot of Confidence Interval Sizes: Bootstrap vs. Standard OLS Output",
+    x = "Confidence Interval Size",
+    y = "Density"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(size = 14, face = "bold"),
+    axis.title = element_text(size = 12),
+    legend.title = element_blank()
+  )
+
+# Violin plot
+ggplot(ci_data_ols, aes(x = Method, y = CI_Size)) +
+  geom_violin(fill = "lightblue") +
+  geom_boxplot(width = 0.1, fill = "white") +
+  theme_minimal() +
+  labs(title = "Distribution of CI_Size by Group", x = "Method", y = "CI_Size")
+
+```
+
+```{r 2c Fit to test data Ridge}
+
+# manually refit final ridge model
+
+# Refit the model manually using glmnet
+X_train <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]
+y_train <- df_pred$Log_Revenue_Total
+
+final_ridge_model <- glmnet(X_train, y_train, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
+
+# Define the columns to include in df_pred
+column_names <- c(
+  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Sources.Purchased",
+  "Sources.Other", "Uses.Retail", "Uses.Resale", "Uses.No.Charge", "Uses.Consumed", "Uses.Losses",
+  "Retail.Residential.Customers", "Retail.Commercial.Customers", "Retail.Industrial.Customers",
+  "Retail.Transportation.Customers", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type"
+)
+
+# save a copy of the holdout dataset to be transformed/cleaned 
+# Create df_test with only the specified columns from holdout
+df_test <- holdout[, column_names]
+
+# Log-transform the Revenue.Total column in the test set to match training
+df_test$Log_Revenue_Total <- log1p(df_test$Revenue.Total)
+
+# Extract predictors (X) and response variable (y) for the test set
+X_test <- model.matrix(Log_Revenue_Total ~ ., data = df_test)[, -1]
+y_test <- df_test$Log_Revenue_Total
+
+# Refit the ridge model on the test set using the final lambda from training
+ridge_test <- glmnet(X_test, y_test, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
+
+# Extract coefficients from the test set
+ridge_test_coefs <- coef(ridge_test)
+
+# Extract coefficients from training and holdout models
+training_coefs <- as.matrix(coef(final_ridge_model, s = final_lambda_ridge))
+holdout_coefs <- as.matrix(ridge_test_coefs)
+
+# Remove the intercept term (row with name "(Intercept)") from both coefficient matrices
+training_coefs <- training_coefs[rownames(training_coefs) != "(Intercept)", , drop = FALSE]
+holdout_coefs <- holdout_coefs[rownames(holdout_coefs) != "(Intercept)", , drop = FALSE]
+
+# Ensure feature names match between the training and holdout coefficients
+common_features <- intersect(rownames(training_coefs), rownames(holdout_coefs))
+
+# Subset coefficients to only include common features
+training_coefs <- training_coefs[common_features, , drop = FALSE]
+holdout_coefs <- holdout_coefs[common_features, , drop = FALSE]
+
+# Combine coefficients into a comparison data frame
+comparison <- data.frame(
+  Predictor = common_features,
+  Training_Coefs = round(as.numeric(training_coefs), 8),
+  Holdout_Coefs = round(as.numeric(holdout_coefs), 8)
+)
+
+# Sort the comparison data frame in decreasing order of Training_Coefs
+comparison <- comparison[order(-comparison$Training_Coefs), ]
+
+# Set options to avoid scientific notation
+options(scipen = 999)
+
+# Display the comparison
+print("Comparison of coefficients from training and holdout data:")
+print(comparison)
+
+```
+
+```{r 2d bootstrap Ridge}
 
@@
 
-```{r 2c Fit to test data Ridge}
-
-# manually refit final ridge model
-
-# Refit the model manually using glmnet
-X_train <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]
-y_train <- df_pred$Log_Revenue_Total
-
-final_ridge_model <- glmnet(X_train, y_train, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
-
-# Define the columns to include in df_pred
-column_names <- c(
-  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Sources.Purchased",
-  "Sources.Other", "Uses.Retail", "Uses.Resale", "Uses.No.Charge", "Uses.Consumed", "Uses.Losses",
-  "Retail.Residential.Customers", "Retail.Commercial.Customers", "Retail.Industrial.Customers",
-  "Retail.Transportation.Customers", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type"
-)
-
-# save a copy of the holdout dataset to be transformed/cleaned 
-# Create df_test with only the specified columns from holdout
-df_test <- holdout[, column_names]
-
-# Log-transform the Revenue.Total column in the test set to match training
-df_test$Log_Revenue_Total <- log1p(df_test$Revenue.Total)
-
-# Extract predictors (X) and response variable (y) for the test set
-X_test <- model.matrix(Log_Revenue_Total ~ ., data = df_test)[, -1]
-y_test <- df_test$Log_Revenue_Total
-
-# Refit the ridge model on the test set using the final lambda from training
-ridge_test <- glmnet(X_test, y_test, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
-
-# Extract coefficients from the test set
-ridge_test_coefs <- coef(ridge_test)
-
-# Extract coefficients from training and holdout models
-training_coefs <- as.matrix(coef(final_ridge_model, s = final_lambda_ridge))
-holdout_coefs <- as.matrix(ridge_test_coefs)
-
-# Remove the intercept term (row with name "(Intercept)") from both coefficient matrices
-training_coefs <- training_coefs[rownames(training_coefs) != "(Intercept)", , drop = FALSE]
-holdout_coefs <- holdout_coefs[rownames(holdout_coefs) != "(Intercept)", , drop = FALSE]
-
-# Ensure feature names match between the training and holdout coefficients
-common_features <- intersect(rownames(training_coefs), rownames(holdout_coefs))
-
-# Subset coefficients to only include common features
-training_coefs <- training_coefs[common_features, , drop = FALSE]
-holdout_coefs <- holdout_coefs[common_features, , drop = FALSE]
-
-# Combine coefficients into a comparison data frame
-comparison <- data.frame(
-  Predictor = common_features,
-  Training_Coefs = round(as.numeric(training_coefs), 8),
-  Holdout_Coefs = round(as.numeric(holdout_coefs), 8)
-)
-
-# Sort the comparison data frame in decreasing order of Training_Coefs
-comparison <- comparison[order(-comparison$Training_Coefs), ]
-
-# Set options to avoid scientific notation
-options(scipen = 999)
-
-# Display the comparison
-print("Comparison of coefficients from training and holdout data:")
-print(comparison)
-
-```
-
-```{r 2d bootstrap Ridge}
-
-# Function to fit Ridge on bootstrap sample and return coefficients
-bootstrap_ridge <- function(data, indices, lambda) {
-  # Bootstrap sampling
-  boot_sample <- data[indices, ]
-
-  # Create the model matrix from the bootstrap sample
-  X_boot <- model.matrix(Log_Revenue_Total ~ ., data = boot_sample)[, -1]
-  y_boot <- boot_sample$Log_Revenue_Total
-
-  # Ensure that X_boot has the same column names as the original data
-  full_X <- model.matrix(Log_Revenue_Total ~ ., data = data)[, -1]
-  full_coefs <- numeric(ncol(full_X))  # Create a vector for all predictors (set to 0 initially)
-  names(full_coefs) <- colnames(full_X)  # Name it with the original predictor names
-
-  # Fit Ridge model
-  ridge_fit <- glmnet(X_boot, y_boot, alpha = 0, lambda = lambda, standardize = TRUE)
-
-  # Extract coefficients (excluding intercept)
-  coefs <- as.vector(coef(ridge_fit)[-1])  # Exclude intercept
-  coef_names <- colnames(X_boot)  # Names of predictors in this bootstrap sample
-
-  # Assign the coefficients to the correct positions
-  full_coefs[coef_names] <- coefs
-
-  return(full_coefs)
-}
-
-# Set up bootstrap
-set.seed(123)
-n_boot <- 1000  # Number of bootstrap samples
-X_train <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]
-y_train <- df_pred$Log_Revenue_Total
-
-# Perform bootstrap
-bootstrap_results <- boot(
-  data = df_pred,
-  statistic = bootstrap_ridge,
-  R = n_boot,
-  lambda = final_lambda_ridge
-)
-
-# Compute confidence intervals for each coefficient
-bootstrap_cis <- apply(bootstrap_results$t, 2, function(coefs) {
-  quantile(coefs, probs = c(0.025, 0.975))
-})
-
-# Format as a data frame
-cis_df <- data.frame(
-  Predictor = colnames(X_train),
-  CI_Lower = bootstrap_cis[1, ],
-  CI_Upper = bootstrap_cis[2, ]
-)
-
-## Compare to the confidence intervals computed in the standard regression output
-
-# Identify numeric predictors
-numeric_predictors <- sapply(df_pred[column_names], is.numeric)
-
-# Standardize only numeric predictors
-df_pred_standardized <- df_pred
-df_pred_standardized[names(numeric_predictors)[numeric_predictors]] <- scale(
-  df_pred[names(numeric_predictors)[numeric_predictors]]
-)
-
-# # See that non-numeric columns remain unchanged
-# print("Standardized dataset:")
-# print(head(df_pred_standardized))
-
-# Fit OLS regression with standardized predictors
-ols_model_standardized <- lm(Log_Revenue_Total ~ ., data = df_pred_standardized)
-
-# Extract OLS confidence intervals
-ols_cis_standardized <- confint(ols_model_standardized, level = 0.95)
-ols_cis_df_standardized <- data.frame(
-  Predictor = rownames(ols_cis_standardized),
-  OLS_CI_Lower_Standardized = ols_cis_standardized[, 1],
-  OLS_CI_Upper_Standardized = ols_cis_standardized[, 2]
-)
-
-# Merge with Bootstrap confidence intervals for comparison
-comparison_ridge_OLS <- merge(
-  cis_df,  # Bootstrap confidence intervals from Ridge
-  ols_cis_df_standardized,
-  by = "Predictor",
-  all.x = TRUE
-)
-
-# Display the comparison
-print("Comparison of Bootstrap and Standardized OLS Confidence Intervals:")
-print(comparison_ridge_OLS)
-
-# visualize the difference in Ridge vs. OLS CIs
-
-# Calculate confidence interval sizes for Ridge and OLS
-comparison_ridge_OLS$Ridge_CI_Size <- comparison_ridge_OLS$CI_Upper - comparison_ridge_OLS$CI_Lower
-comparison_ridge_OLS$OLS_CI_Size <- comparison_ridge_OLS$OLS_CI_Upper_Standardized - comparison_ridge_OLS$OLS_CI_Lower_Standardized
-
-# Reshape the data for plotting
-ci_data_ridge <- comparison_ridge_OLS %>%
-  select(Predictor, Ridge_CI_Size, OLS_CI_Size) %>%
-  pivot_longer(cols = c(Ridge_CI_Size, OLS_CI_Size), 
-               names_to = "Method", 
-               values_to = "CI_Size")
-
-# Rename methods for better readability
-ci_data_ridge$Method <- ifelse(ci_data_ridge$Method == "Ridge_CI_Size", "Ridge", "OLS")
-
-# Create the density plot
-ggplot(ci_data_ridge, aes(x = CI_Size, fill = Method)) +
-  geom_density(alpha = 0.5) +
-  scale_x_log10() +
-  labs(
-    title = "Density Plot of Confidence Interval Sizes: Ridge vs. OLS",
-    x = "Confidence Interval Size",
-    y = "Density"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(size = 14, face = "bold"),
-    axis.title = element_text(size = 12),
-    legend.title = element_blank()
-  )
-
-# violin plot
-ggplot(ci_data_ridge, aes(x = Method, y = CI_Size)) +
-  geom_violin(fill = "lightblue") +
-  geom_boxplot(width = 0.1, fill = "white") +
-  theme_minimal() +
-  labs(title = "Distribution of CI_Size by Group", x = "Group", y = "CI_Size")
-
-```
-
 ```{r 2c Fit to test data Lasso}

### 2024-12-03 MS&E 226_Project Part 2_Code.Rmd
2e multiple hypothesis testing

@@
 
+#### (d)
+
 ```{r 2d bootstrap OLS}
@@
 
+#### (e)
+
+```{r Benjamini_Hochberg Procedure}
+
+# Extract the p-values for the coefficients
+p_values <- summary(ols_test)$coefficients[, 4] # 4th column contains p-values
+
+# Apply the Benjamini-Hochberg procedure
+adjusted_p_values <- p.adjust(p_values, method = "BH")
+
+# Combine the results into a summary dataframe
+results <- data.frame(
+  Coefficient = rownames(summary(ols_test)$coefficients),
+  P_Value = p_values,
+  Adjusted_P_Value = adjusted_p_values
+)
+
+# Display the results
+print(results)
+
+# Optional: Identify which hypotheses are significant after FDR correction
+significant <- results$Adjusted_P_Value < 0.05
+results$Significant <- ifelse(significant, "Yes", "No")
+
+# Display the results with significance labels
+print(results)
+
+```
+
+#### GRAVEYARD
+
 ```{r 2c Fit to test data Ridge}

### 2024-12-04 MS&E 226_Project Part 2_Code.Rmd
ols significant coeff

@@
 
-#### (a)
+### Statistically Significant Coefficients
 ```{r 2a}
-# Extract the final Ridge model using the best lambda
-final_ridge_model <- ridge_model$finalModel
-
-# Extract coefficients from the final model
-coefficients <- coef(final_ridge_model, s = final_lambda_ridge)
-
-# Print the coefficients
-print(coefficients)
-
-```
-
-#### (b)
+# Display summary of the OLS model
+summary_ols <- summary(ols_train)
+
+# Extract coefficients table
+coefficients_table <- summary_ols$coefficients
+
+# Identify statistically significant coefficients
+significant_coefficients <- coefficients_table[coefficients_table[, 4] < 0.05, ]
+
+# Print significant coefficients
+print(significant_coefficients)
+
+```
+
+### Subset Model
 ```{r}
 # Select a subset of covariates
-subset_column_names <- c(  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type")
-
-# Create a new data frame with the subset of columns
+# Subset the dataframe
+subset_column_names <- c("Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", 
+                         "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type")
 df_subset <- df[, subset_column_names]
-
 df_subset$Log_Revenue_Total <- log1p(df_subset$Revenue.Total)
 
-# Train the Ridge Regression model using the subset of covariates
-ridge_subset_model <- train(Log_Revenue_Total ~ ., 
-                            data = df_subset, 
-                            method = "glmnet",  
-                            trControl = train_control,
-                            metric = "RMSE", 
-                            preProcess = "scale",  
-                            tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 1, length = 100)))
-
-```
+# Create the formula for the subset model
+subset_formula <- as.formula("Log_Revenue_Total ~ Demand.Summer.Peak + Demand.Winter.Peak + 
+                              Sources.Generation + Utility.Region + Max.Revenue.Sector + Utility.Type")
+
+# Fit the subset OLS model
+subset_ols <- lm(subset_formula, data = df_subset)
+
+# Print the summary of the subset model
+summary(subset_ols)
+
+```
+
 ```{r}
-# Get the coefficients for the best model at the selected lambda
-best_lambda <- ridge_model$bestTune$lambda
-best_model_coeffs <- coef(ridge_model$finalModel, s = best_lambda)
-
-# Print all coefficients for the best Ridge Regression model
-cat("All coefficients for the best Ridge Regression model:\n")
-print(best_model_coeffs)
-
-# Get coefficients for the subset model
-best_lambda_subset <- ridge_subset_model$bestTune$lambda
-best_model_coeffs_subset <- coef(ridge_subset_model$finalModel, s = best_lambda_subset)
-
-# Print all coefficients for the subset model
-cat("All coefficients for the Ridge Regression model with a subset of covariates:\n")
-print(best_model_coeffs_subset)
-```
-
-#### (c)
+# Extract significant coefficients from the subset model
+subset_summary <- summary(subset_ols)
+subset_coefficients <- subset_summary$coefficients
+significant_subset_coefficients <- subset_coefficients[subset_coefficients[, 4] < 0.05, ]
+
+# Print significant coefficients from the subset model
+print(significant_subset_coefficients)
+```
+
+### OLS with test data
 
@@
 # Combine p-values into a comparison data frame
-comparison <- data.frame(
+comparison_pvalues <- data.frame(
   Predictor = common_features,
@@
 # Sort the comparison data frame in increasing order of Training_Pvalues
-comparison <- comparison[order(comparison$Training_Pvalues), ]
+comparison_pvalues <- comparison_pvalues[order(comparison_pvalues$Training_Pvalues), ]
 
@@
 
-# Display the comparison
+# Display the comparison of p-values
 print("Comparison of p-values from training and test data:")
-print(comparison)
-
-## compare coefficients
+print(comparison_pvalues)
+
+# Extract coefficients from the training and test models
+training_coefs <- coef(ols_train)
+holdout_coefs <- coef(ols_test)
+
+# Ensure feature names match between training and holdout coefficients
+common_features <- intersect(names(training_coefs), names(holdout_coefs))
 
 # Subset coefficients to only include common features
-training_coefs <- training_coefs[common_features, , drop = FALSE]
-holdout_coefs <- holdout_coefs[common_features, , drop = FALSE]
+training_coefs <- training_coefs[common_features]
+holdout_coefs <- holdout_coefs[common_features]
 
 # Combine coefficients into a comparison data frame
-comparison <- data.frame(
+comparison_coefs <- data.frame(
   Predictor = common_features,
@@
 # Sort the comparison data frame in decreasing order of Training_Coefs
-comparison <- comparison[order(-comparison$Training_Coefs), ]
-
-# Display the comparison
+comparison_coefs <- comparison_coefs[order(-comparison_coefs$Training_Coefs), ]
+
+# Display the comparison of coefficients
 print("Comparison of coefficients from training and holdout data:")
-print(comparison)
-
-```
-
-#### (d)
+print(comparison_coefs)
+
+```
+
+### Bootstrap
 
@@
 
-#### (e)
+### Multiple Hypothesis Testing
 
@@
 ```
-
-#### GRAVEYARD
-
-```{r 2c Fit to test data Ridge}
-
-# manually refit final ridge model
-
-# Refit the model manually using glmnet
-X_train <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]
-y_train <- df_pred$Log_Revenue_Total
-
-final_ridge_model <- glmnet(X_train, y_train, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
-
-# Define the columns to include in df_pred
-column_names <- c(
-  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Sources.Purchased",
-  "Sources.Other", "Uses.Retail", "Uses.Resale", "Uses.No.Charge", "Uses.Consumed", "Uses.Losses",
-  "Retail.Residential.Customers", "Retail.Commercial.Customers", "Retail.Industrial.Customers",
-  "Retail.Transportation.Customers", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type"
-)
-
-# save a copy of the holdout dataset to be transformed/cleaned 
-# Create df_test with only the specified columns from holdout
-df_test <- holdout[, column_names]
-
-# Log-transform the Revenue.Total column in the test set to match training
-df_test$Log_Revenue_Total <- log1p(df_test$Revenue.Total)
-
-# Extract predictors (X) and response variable (y) for the test set
-X_test <- model.matrix(Log_Revenue_Total ~ ., data = df_test)[, -1]
-y_test <- df_test$Log_Revenue_Total
-
-# Refit the ridge model on the test set using the final lambda from training
-ridge_test <- glmnet(X_test, y_test, alpha = 0, lambda = final_lambda_ridge, standardize = TRUE)
-
-# Extract coefficients from the test set
-ridge_test_coefs <- coef(ridge_test)
-
-# Extract coefficients from training and holdout models
-training_coefs <- as.matrix(coef(final_ridge_model, s = final_lambda_ridge))
-holdout_coefs <- as.matrix(ridge_test_coefs)
-
-# Remove the intercept term (row with name "(Intercept)") from both coefficient matrices
-training_coefs <- training_coefs[rownames(training_coefs) != "(Intercept)", , drop = FALSE]
-holdout_coefs <- holdout_coefs[rownames(holdout_coefs) != "(Intercept)", , drop = FALSE]
-
-# Ensure feature names match between the training and holdout coefficients
-common_features <- intersect(rownames(training_coefs), rownames(holdout_coefs))
-
-# Subset coefficients to only include common features
-training_coefs <- training_coefs[common_features, , drop = FALSE]
-holdout_coefs <- holdout_coefs[common_features, , drop = FALSE]
-
-# Combine coefficients into a comparison data frame
-comparison <- data.frame(
-  Predictor = common_features,
-  Training_Coefs = round(as.numeric(training_coefs), 8),
-  Holdout_Coefs = round(as.numeric(holdout_coefs), 8)
-)
-
-# Sort the comparison data frame in decreasing order of Training_Coefs
-comparison <- comparison[order(-comparison$Training_Coefs), ]
-
-# Set options to avoid scientific notation
-options(scipen = 999)
-
-# Display the comparison
-print("Comparison of coefficients from training and holdout data:")
-print(comparison)
-
-```
-
-```{r 2d bootstrap Ridge}
-
-# Function to fit Ridge on bootstrap sample and return coefficients
-bootstrap_ridge <- function(data, indices, lambda) {
-  # Bootstrap sampling
-  boot_sample <- data[indices, ]
-
-  # Create the model matrix from the bootstrap sample
-  X_boot <- model.matrix(Log_Revenue_Total ~ ., data = boot_sample)[, -1]
-  y_boot <- boot_sample$Log_Revenue_Total
-
-  # Ensure that X_boot has the same column names as the original data
-  full_X <- model.matrix(Log_Revenue_Total ~ ., data = data)[, -1]
-  full_coefs <- numeric(ncol(full_X))  # Create a vector for all predictors (set to 0 initially)
-  names(full_coefs) <- colnames(full_X)  # Name it with the original predictor names
-
-  # Fit Ridge model
-  ridge_fit <- glmnet(X_boot, y_boot, alpha = 0, lambda = lambda, standardize = TRUE)
-
-  # Extract coefficients (excluding intercept)
-  coefs <- as.vector(coef(ridge_fit)[-1])  # Exclude intercept
-  coef_names <- colnames(X_boot)  # Names of predictors in this bootstrap sample
-
-  # Assign the coefficients to the correct positions
-  full_coefs[coef_names] <- coefs
-
-  return(full_coefs)
-}
-
-# Set up bootstrap
-set.seed(123)
-n_boot <- 1000  # Number of bootstrap samples
-X_train <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]
-y_train <- df_pred$Log_Revenue_Total
-
-# Perform bootstrap
-bootstrap_results <- boot(
-  data = df_pred,
-  statistic = bootstrap_ridge,
-  R = n_boot,
-  lambda = final_lambda_ridge
-)
-
-# Compute confidence intervals for each coefficient
-bootstrap_cis <- apply(bootstrap_results$t, 2, function(coefs) {
-  quantile(coefs, probs = c(0.025, 0.975))
-})
-
-# Format as a data frame
-cis_df <- data.frame(
-  Predictor = colnames(X_train),
-  CI_Lower = bootstrap_cis[1, ],
-  CI_Upper = bootstrap_cis[2, ]
-)
-
-## Compare to the confidence intervals computed in the standard regression output
-
-# Identify numeric predictors
-numeric_predictors <- sapply(df_pred[column_names], is.numeric)
-
-# Standardize only numeric predictors
-df_pred_standardized <- df_pred
-df_pred_standardized[names(numeric_predictors)[numeric_predictors]] <- scale(
-  df_pred[names(numeric_predictors)[numeric_predictors]]
-)
-
-# # See that non-numeric columns remain unchanged
-# print("Standardized dataset:")
-# print(head(df_pred_standardized))
-
-# Fit OLS regression with standardized predictors
-ols_model_standardized <- lm(Log_Revenue_Total ~ ., data = df_pred_standardized)
-
-# Extract OLS confidence intervals
-ols_cis_standardized <- confint(ols_model_standardized, level = 0.95)
-ols_cis_df_standardized <- data.frame(
-  Predictor = rownames(ols_cis_standardized),
-  OLS_CI_Lower_Standardized = ols_cis_standardized[, 1],
-  OLS_CI_Upper_Standardized = ols_cis_standardized[, 2]
-)
-
-# Merge with Bootstrap confidence intervals for comparison
-comparison_ridge_OLS <- merge(
-  cis_df,  # Bootstrap confidence intervals from Ridge
-  ols_cis_df_standardized,
-  by = "Predictor",
-  all.x = TRUE
-)
-
-# Display the comparison
-print("Comparison of Bootstrap and Standardized OLS Confidence Intervals:")
-print(comparison_ridge_OLS)
-
-# visualize the difference in Ridge vs. OLS CIs
-
-# Calculate confidence interval sizes for Ridge and OLS
-comparison_ridge_OLS$Ridge_CI_Size <- comparison_ridge_OLS$CI_Upper - comparison_ridge_OLS$CI_Lower
-comparison_ridge_OLS$OLS_CI_Size <- comparison_ridge_OLS$OLS_CI_Upper_Standardized - comparison_ridge_OLS$OLS_CI_Lower_Standardized
-
-# Reshape the data for plotting
-ci_data_ridge <- comparison_ridge_OLS %>%
-  select(Predictor, Ridge_CI_Size, OLS_CI_Size) %>%
-  pivot_longer(cols = c(Ridge_CI_Size, OLS_CI_Size), 
-               names_to = "Method", 
-               values_to = "CI_Size")
-
-# Rename methods for better readability
-ci_data_ridge$Method <- ifelse(ci_data_ridge$Method == "Ridge_CI_Size", "Ridge", "OLS")
-
-# Create the density plot
-ggplot(ci_data_ridge, aes(x = CI_Size, fill = Method)) +
-  geom_density(alpha = 0.5) +
-  scale_x_log10() +
-  labs(
-    title = "Density Plot of Confidence Interval Sizes: Ridge vs. OLS",
-    x = "Confidence Interval Size",
-    y = "Density"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(size = 14, face = "bold"),
-    axis.title = element_text(size = 12),
-    legend.title = element_blank()
-  )
-
-# violin plot
-ggplot(ci_data_ridge, aes(x = Method, y = CI_Size)) +
-  geom_violin(fill = "lightblue") +
-  geom_boxplot(width = 0.1, fill = "white") +
-  theme_minimal() +
-  labs(title = "Distribution of CI_Size by Group", x = "Group", y = "CI_Size")
-
-```
-
-```{r 2c Fit to test data Lasso}
-
-# Define the columns to include in df_pred
-column_names <- c(
-  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Sources.Purchased",
-  "Sources.Other", "Uses.Retail", "Uses.Resale", "Uses.No.Charge", "Uses.Consumed", "Uses.Losses",
-  "Retail.Residential.Customers", "Retail.Commercial.Customers", "Retail.Industrial.Customers",
-  "Retail.Transportation.Customers", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type"
-)
-
-# save a copy of the holdout dataset to be transformed/cleaned 
-# Create df_test with only the specified columns from holdout
-df_test <- holdout[, column_names]
-
-# Log-transform the Revenue.Total column in the test set to match training
-df_test$Log_Revenue_Total <- log1p(df_test$Revenue.Total)
-
-# # Extract predictors (X) and response variable (y) for the test set
-# X_test <- model.matrix(Log_Revenue_Total ~ ., data = df_test)[, -1]
-# y_test <- df_test$Log_Revenue_Total
-
-# Refit the Lasso model on the test set using the final lambda from training
-lasso_test <- glmnet(X_test, y_test, alpha = 1, lambda = final_lambda_lasso, standardize = TRUE)
-
-# Extract coefficients from the test set
-lasso_test_coefs <- coef(lasso_test)
-
-# Extract coefficients from training and holdout models
-training_coefs <- as.matrix(coef(lasso_model$finalModel, s = final_lambda_lasso))
-holdout_coefs <- as.matrix(lasso_test_coefs)
-
-# Remove the intercept term (row with name "(Intercept)") from both coefficient matrices
-training_coefs <- training_coefs[rownames(training_coefs) != "(Intercept)", , drop = FALSE]
-holdout_coefs <- holdout_coefs[rownames(holdout_coefs) != "(Intercept)", , drop = FALSE]
-
-# Ensure feature names match between the training and holdout coefficients
-common_features <- intersect(rownames(training_coefs), rownames(holdout_coefs))
-
-# Subset coefficients to only include common features
-training_coefs <- training_coefs[common_features, , drop = FALSE]
-holdout_coefs <- holdout_coefs[common_features, , drop = FALSE]
-
-# Combine coefficients into a comparison data frame
-comparison <- data.frame(
-  Predictor = common_features,
-  Training_Coefs = round(as.numeric(training_coefs), 8),
-  Holdout_Coefs = round(as.numeric(holdout_coefs), 8)
-)
-
-# Sort the comparison data frame in decreasing order of Training_Coefs
-comparison <- comparison[order(-comparison$Training_Coefs), ]
-
-# Set options to avoid scientific notation
-options(scipen = 999)
-
-# Display the comparison
-print("Comparison of coefficients from training and holdout data:")
-print(comparison)
-
-```
-
-```{r 2d bootstrap Lasso}
-
-# Function to fit Lasso on bootstrap sample and return coefficients
-bootstrap_lasso <- function(data, indices, lambda) {
-  # Bootstrap sampling
-  boot_sample <- data[indices, ]
-
-  # Create the model matrix from the bootstrap sample
-  X_boot <- model.matrix(Log_Revenue_Total ~ ., data = boot_sample)[, -1]
-  y_boot <- boot_sample$Log_Revenue_Total
-
-  # Ensure that X_boot has the same column names as the original data
-  full_X <- model.matrix(Log_Revenue_Total ~ ., data = data)[, -1]
-  full_coefs <- numeric(ncol(full_X))  # Create a vector for all predictors (set to 0 initially)
-  names(full_coefs) <- colnames(full_X)  # Name it with the original predictor names
-
-  # Fit Lasso model
-  lasso_fit <- glmnet(X_boot, y_boot, alpha = 1, lambda = lambda, standardize = TRUE)
-
-  # Extract coefficients (excluding intercept)
-  coefs <- as.vector(coef(lasso_fit)[-1])  # Exclude intercept
-  coef_names <- colnames(X_boot)  # Names of predictors in this bootstrap sample
-
-  # Assign the coefficients to the correct positions
-  full_coefs[coef_names] <- coefs
-
-  return(full_coefs)
-}
-
-# Set up bootstrap
-set.seed(123)
-n_boot <- 1000  # Number of bootstrap samples
-X_train <- model.matrix(Log_Revenue_Total ~ ., data = df_pred)[, -1]
-y_train <- df_pred$Log_Revenue_Total
-
-# Perform bootstrap
-bootstrap_results <- boot(
-  data = df_pred,
-  statistic = bootstrap_lasso,
-  R = n_boot,
-  lambda = final_lambda_lasso
-)
-
-# Compute confidence intervals for each coefficient
-bootstrap_cis <- apply(bootstrap_results$t, 2, function(coefs) {
-  quantile(coefs, probs = c(0.025, 0.975))
-})
-
-# Format as a data frame
-cis_df <- data.frame(
-  Predictor = colnames(X_train),
-  CI_Lower = bootstrap_cis[1, ],
-  CI_Upper = bootstrap_cis[2, ]
-)
-
-## Compare to the confidence intervals computed in the standard regression output
-
-# Identify numeric predictors
-numeric_predictors <- sapply(df_pred[column_names], is.numeric)
-
-# Standardize only numeric predictors
-df_pred_standardized <- df_pred
-df_pred_standardized[names(numeric_predictors)[numeric_predictors]] <- scale(
-  df_pred[names(numeric_predictors)[numeric_predictors]]
-)
-
-# # See that non-numeric columns remain unchanged
-# print("Standardized dataset:")
-# print(head(df_pred_standardized))
-
-# Fit OLS regression with standardized predictors
-ols_model_standardized <- lm(Log_Revenue_Total ~ ., data = df_pred_standardized)
-
-# Extract OLS confidence intervals
-ols_cis_standardized <- confint(ols_model_standardized, level = 0.95)
-ols_cis_df_standardized <- data.frame(
-  Predictor = rownames(ols_cis_standardized),
-  OLS_CI_Lower_Standardized = ols_cis_standardized[, 1],
-  OLS_CI_Upper_Standardized = ols_cis_standardized[, 2]
-)
-
-# Merge with Bootstrap confidence intervals for comparison
-comparison_lasso_OLS <- merge(
-  cis_df,  # Bootstrap confidence intervals from Lasso
-  ols_cis_df_standardized,
-  by = "Predictor",
-  all.x = TRUE
-)
-
-# Display the comparison
-print("Comparison of Bootstrap and Standardized OLS Confidence Intervals:")
-print(comparison_lasso_OLS)
-
-# visualize the difference in Lasso vs. OLS CIs
-
-# Calculate confidence interval sizes for Lasso and OLS
-comparison_lasso_OLS$Lasso_CI_Size <- comparison_lasso_OLS$CI_Upper - comparison_lasso_OLS$CI_Lower
-comparison_lasso_OLS$OLS_CI_Size <- comparison_lasso_OLS$OLS_CI_Upper_Standardized - comparison_lasso_OLS$OLS_CI_Lower_Standardized
-
-# Reshape the data for plotting
-ci_data_lasso <- comparison_lasso_OLS %>%
-  select(Predictor, Lasso_CI_Size, OLS_CI_Size) %>%
-  pivot_longer(cols = c(Lasso_CI_Size, OLS_CI_Size), 
-               names_to = "Method", 
-               values_to = "CI_Size")
-
-# Rename methods for better readability
-ci_data_lasso$Method <- ifelse(ci_data_lasso$Method == "Lasso_CI_Size", "Lasso", "OLS")
-
-# Create the density plot
-ggplot(ci_data_lasso, aes(x = CI_Size, fill = Method)) +
-  geom_density(alpha = 0.5) +
-  scale_x_log10() +
-  labs(
-    title = "Density Plot of Confidence Interval Sizes: Lasso vs. OLS",
-    x = "Confidence Interval Size",
-    y = "Density"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(size = 14, face = "bold"),
-    axis.title = element_text(size = 12),
-    legend.title = element_blank()
-  )
-
-# violin plot
-ggplot(ci_data_lasso, aes(x = Method, y = CI_Size)) +
-  geom_violin(fill = "lightblue") +
-  geom_boxplot(width = 0.1, fill = "white") +
-  theme_minimal() +
-  labs(title = "Distribution of CI_Size by Group", x = "Group", y = "CI_Size")
-
-```

### 2024-12-04 MS&E 226_Project Part 2_Code.Rmd
final edits

@@
 
-# Print significant coefficients
-print(significant_coefficients)
+significant_coefficients_df <- data.frame(
+  Variable = rownames(significant_coefficients),
+  Pr_greater_than_t = significant_coefficients[, 4]
+)
+
+# Print the table
+print(significant_coefficients_df)
 
@@
 
-# Print significant coefficients from the subset model
-print(significant_subset_coefficients)
+significant_subset_coefficients_df <- data.frame(
+  Variable = rownames(significant_subset_coefficients),
+  Pr_greater_than_t = significant_subset_coefficients[, 4]
+)
+
+# Print the table
+print(significant_subset_coefficients_df)
 ```
